---
title: "0100_Data_from_excel_2020"
author: "DHJ"
date: "2022-11-11"
output: html_document
---


No data in Vannmiljø for 2016-2019

In https://vannmiljo.miljodirektoratet.no/
"Jeg vil" => Søk i vannregistreringer og lokaliteter
Select records: 
  in menu "Forvaltningsaktivitet": select "Miljøgifter i en urban fjord"
  in menu "Prøvedato": select from 01/01/2016 to 31/12/2019
Returns no records ("Info! Søket gav ingen treff")

---------------------------------------------------------------

PARAMETERS
(see part 8)

NIVA:
- PFAS
- UV-chemicals (Octocrylene, benzophenone-3, ethylhexylmethoxycinnamate (plus UV-
    327, -328 and -329))
- Behentrimonium (ATAC-C20 and ATAC-C22)
- Biological effect parameters: Acetylcholinesterase (AChE)

NILU:
- metals
- PCBs, BDEs, brominated flame retardants and CPs (SCCP, MCCP)  
- alkylphenols and bisphenols
- siloxans
- PFRs (phosphorous flame retardants)
- antioxidant MB1
- M3T(Ph)
- Dechlorane plus and related compounds

IFE:
- stable isotopes of carbon and nitrogen

The concentrations of the following contaminants were expressed on a lipid weight basis: 
 PCBs and other organochlorine compounds, chlorinated paraffins, brominated flame retardants, 
 siloxanes (including M3T(Ph)), UV-filters and declorane plus (antioxidant MB1 was not 
 analysed in sufficient species for such evaluation in 2019). 
Behentrimonium was expressed at both wet weight and lipid weight basis
 when exploring correlations between contaminant concentrations and trophic position.

TABLES OF SUBSTANCES USED  
df_substances  
- contains substance names (used in Excel), Decription (long name), Chemspider and CAS   
- this equals appendix table B1 of the report  
df_substance_lookup  
- contains substance names (used in Excel), Group and Lab   
  

CHECKLIST FOR IMPORT - WHEN DATA ARE IN LABWARE_IMPORT + LABWARE_CHECK_SAMPLE, BUT NOT IN NIVADATABASE:  
1. Has STATION_CODE been set?  
2. Has sample number (e.g. 1-15) been set?  
  
If not 1 and/or 2, the lab has to make update!  
  

## 1. Setup  

### Functions  
```{r}

# library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)
library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```

### Year  
```{r}

year_sampling <- 2020

```



### Nivadatabase password   
* Not needed - assume that password is set ising the 'keyring' package  
```{r, echo = FALSE}

### Set Nivadabase password 

# Do this once:
# key_set_with_value(service = "NIVAbasen",
#                    username = "DHJ",
#                    password = "***")

# db_username <- "DHJ"
# db_privkey <- sodium::keygen()
# db_pubkey <- sodium::pubkey(db_privkey)
# db_pwd <- sodium::simple_encrypt(
#   serialize(
#     keyring::key_get(service = "NIVAbasen", username = "DHJ"), NULL), 
#   db_pubkey)

# Or do this:
# set_credentials()

    
```


### Lookup tables   
* Substance to substance group (Group) + Lab (NIVA / NILU etc)     
* Substance to substance name and CAS (used for populating METHOD_DEFINITIONS)     
* TISSUE_ID to TISSUE   
* SPECIE_ID to LATIN_NAME  
```{r}

# Look-up for substance groups (Group) + Lab (NIVA / NILU etc)    
# "Hand-made" in Excel
lookup_substance_groups <- readxl::read_excel("Input_data/Lookup_files/992_Excel_substances.xlsx")

# Look-up for substance names + CAS numbers   
# - contains substance names (used in Excel), Decription (long name), Chemspider and CAS 
# - this equals appendix table B1 of the report
lookup_substance_names <- readRDS("Input_data/Lookup_files/Report2019_substancetable.rds") 

# From Nivabasen
# TISSUE_ID 

if (FALSE){
  # If we want to read from Nivabase
  lookup_tissues <- get_nivabase_data("select * from LABWARE.V_NIVA_TISSUE_TYPES", source = 'NIVABASE')
  lookup_species <- get_nivabase_data("select * from LABWARE.V_NIVA_SPECIES", source = 'NIVABASE')
}

lookup_tissues <- read.csv("Input_data/Lookup_files/lookup_tissues.csv")
lookup_species <- read.csv("Input_data/Lookup_files/lookup_species.csv")

```


## 2. Project ID  
```{r}

df_projects <- get_projects() 
# Add O-numbers
df_o_numbers <- get_nivabase_data("select PROJECT_ID, O_NUMBER from NIVADATABASE.PROJECTS_O_NUMBERS") %>%
  group_by(PROJECT_ID) %>%
  summarise(O_NUMBER = paste(O_NUMBER, collapse = ","), .groups = "drop")
df_projects <- left_join(df_projects, df_o_numbers, by = "PROJECT_ID")

# Pick projects by name
df_projects_sel <- df_projects %>% 
  filter(grepl("urban", PROJECT_NAME, ignore.case = TRUE))

proj_ids <- unique(df_projects_sel$PROJECT_ID)

df_projects_sel

```

## 3. Check existing samples   

### Get stations and station IDs from old data       
```{r}

df_stations <- readRDS("Input_data/Lookup_files/992_df_stations_2019.rds") %>%
  ungroup() %>%
  rename(STATION_ID = AQUAMONITOR_ID)

df_station_names <- get_nivabase_selection(
  "STATION_ID, STATION_CODE, STATION_NAME, PROJECT_ID",
  "PROJECTS_STATIONS",
  "STATION_ID",
  df_stations$STATION_ID, 
  extra_sql = "and PROJECTS_STATIONS.PROJECT_ID = 8551"
)

df_stations <- df_stations %>%
  left_join(df_station_names)
  
writexl::write_xlsx(df_stations, "Data/lookup_stations_urbanfjord.xlsx")


```

### Add station type    
```{r}

df_for_join1 <- get_nivabase_selection(
  "STATION_ID, STATION_TYPE_ID",
  "STATIONS",
  "STATION_ID",
  df_stations$STATION_ID)

df_for_join2 <- get_nivabase_selection(
  "STATION_TYPE_ID, STATION_TYPE, STATION_TYPE_DESCR",
  "STATION_TYPES",
  "STATION_TYPE_ID",
  unique(df_for_join1$STATION_TYPE_ID))

df_stations <- df_stations %>%
  left_join(df_for_join1) %>%
  left_join(df_for_join2)

df_stations %>%
  group_by(STATION_TYPE) %>%
  summarise(AQUAMONITOR_CODEs = paste(AQUAMONITOR_CODE, collapse = ";"))

```

### Get biota measurements, 2020  
```{r}

df2 <- get_nivabase_selection(
  "STATION_ID, SPECIMEN_ID, SPECIMEN_NO, DATE_CAUGHT, TAXONOMY_CODE_ID",
  "BIOTA_SINGLE_SPECIMENS",
  "STATION_ID",
  unique(df_stations$STATION_ID)  
)

unique(year(df2$DATE_CAUGHT))  

df3 <- get_nivabase_selection(
  "SPECIMEN_ID, SAMPLE_ID",
  "BIOTA_SAMPLES_SPECIMENS",
  "SPECIMEN_ID",
  unique(subset(df2, year(df2$DATE_CAUGHT) == year_sampling)$SPECIMEN_ID)  
)

# Standard units (from script 994-2020)
fn <- "../Milkys2_pc/Files_to_Jupyterhub_2019/Lookup table - standard units.xlsx"
lookup_standard_units <- read_excel(fn)

# IMPORTANT! Add MG_P_KG to MG_P_KG!
lookup_standard_units <- lookup_standard_units %>%
  bind_rows(
    tibble(UNIT_orig = "MG_P_KG", UNIT = "MG_P_KG", Factor = 1)
  )

# 2 minutes  
# debugonce(get_biota_chemistry)
dat_biota <- get_biota_chemistry(
  year_sampling,
  specimendata = df2,
  stationdata = df_stations
)

```

#### Water 2020  

```{r}

# Get projects with data from Skjælholmene
id <- df_stations %>% filter(STATION_NAME == "Skjælholmene") %>% pull(STATION_ID) 

df_proj_stat1 <- get_nivabase_selection(
  "PROJECT_ID, STATION_ID",
  "PROJECTS_STATIONS",
  "STATION_ID",
  id
)
# Only one project ID: 8551

# Get all station IDs
df_proj_stat2 <- get_nivabase_selection(
  "PROJECT_ID, STATION_ID",
  "PROJECTS_STATIONS",
  "PROJECT_ID",
  8551
)

dat_water <- get_water_chemistry_from_stations(unique(df_proj_stat2$STATION_ID), year_sampling)


```


#### Sediment 2020  

```{r}

dat_sediment <- get_sediment_chemistry_from_stations(unique(df_proj_stat2$STATION_ID), year_sampling)

dat_sediment_methods <- get_chemistry_methods(method_id = unique(dat_sediment$METHOD_ID))

dat_sediment <- dat_sediment %>%
  left_join(dat_sediment_methods)

```

## 4. Join with with labware data  

### All labware sample data for these stations   

* Note: all years!

```{r}

df_samples_labware_initial <- get_nivabase_selection(
    "TEXT_ID, STATUS, SPECIES, TISSUE, BIOTA_SAMPLENO, X_BULK_BIO, DESCRIPTION, AQUAMONITOR_ID, AQUAMONITOR_CODE", 
    "LABWARE_CHECK_SAMPLE",
    "AQUAMONITOR_ID", 
    df_stations$STATION_ID, values_are_text = FALSE)

nrow(df_samples_labware_initial)
# 435

df_samples_labware_initial$TEXT_ID %>% substring(4,7) %>% table()

# df_samples_labware_initial$TEXT_ID %>% table() %>% table()

```


### Get data from LABWARE_IMPORT and LABWARE_CHECK_SAMPLE    
Using function 'get_labware_samples'  
```{r}

df_samples_labware <- get_labware_samples(df_samples_labware_initial$TEXT_ID)

```



### Show summary of sample numbers  
```{r}

df1 <- df_samples_labware %>%
  mutate(ID = as.numeric(substr(TEXT_ID, 9, 13)),
         Year = as.numeric(substr(TEXT_ID, 4, 7))) %>%
  group_by(SAMPLE_TYPE, AQUAMONITOR_ID, AQUAMONITOR_CODE, TISSUE, Year) %>%
  summarise(n = n(), Id_min = min(ID), Id_max = max(ID), .groups = "drop") %>%
  arrange(Year, Id_min)

df1

df1 %>% 
  select(SAMPLE_TYPE, AQUAMONITOR_ID, AQUAMONITOR_CODE, TISSUE, Year, n) %>%
  pivot_wider(names_from = Year, values_from = n)

```

## 6. External data    

### a. NILU  


#### Read and select rows    

```{r}

fn <- "Input_data/2020/NILU/004C NILU data stacked.txt"
df_nilu_all <- read.delim(fn)  

if (FALSE){
  
# All data = including rows with no actual data in them
df_nilu_all %>%
  distinct(Matrix, Label) %>%
  xtabs(~Matrix, .)
  
}

# Empty cells are not NA, they are empty character ("")
sel_rows <- df_nilu_all$Data != "" | df_nilu_all$LOQ  != ""

message("Fraction of rows with data: ", mean(sel_rows))  # 0.67

#
# Actual data
#
df_nilu <- df_nilu_all[sel_rows,]

cat(nrow(df_nilu), "rows\n")

```

#### Number of substances  

```{r}

cat("Number of parameters: \n\n") 
df_nilu %>%              # View()
  distinct(Matrix, Label) %>%
  xtabs(~Matrix, .)

```

#### Actual substances  

- Save overview of parameters and ID

```{r}

# A random sample  
df_nilu %>% 
  xtabs(~Label, .) %>% 
  sample(30)

# Save to file
list(
  Parameters = df_nilu %>% count(Label),
  TEXT_ID = df_nilu %>% count(LIMS.NR)
  ) %>%
  writexl::write_xlsx("Data/NILU_2020_overview.xlsx")

```

#### Check NR numbers  

- All TEXT_ID in labware samples are found in NILU data  
- NILU data have 16 bile samples just marked with "galle1-galle16" - phenols measured  

```{r}

nr1 <- df_samples_labware %>% 
  filter(grepl(as.character(year_sampling), TEXT_ID)) %>%
  pull(TEXT_ID)

nr2 <- df_nilu %>%
  distinct(LIMS.NR) %>%
  pull(LIMS.NR)

# setdiff(1:5, 2:3)
setdiff(nr1, nr2)
setdiff(nr2, nr1)  # Galle 1-15 is in nr2 (NILU data) but not in 'df_samples_labware'  

```


### IFE  

- Isotopes  

#### Read and select rows    

```{r}

fn <- "Input_data/2020/IFE/010 LIMS numbers and data IFE.txt"
readLines(fn, 2)

df_ife_all <- read.delim(fn)  

# Empty cells are not NA, they are empty character ("")
sel_rows <- df_ife_all$d13CVPDB != "" 

message("Fraction of rows with data: ", mean(sel_rows))  # 0.67

# View(df_ife_all[!sel_rows,])

#
# Actual data
#
df_ife <- df_ife_all[sel_rows,]

cat(nrow(df_ife), "rows\n")

```

#### Lines with no data has some pooled sample info  
```{r}

df_ife_all[!sel_rows,] %>%
  filter(Column.2 != "") %>%
  select(Column.2, Column.3, Column.4, Individual)

```
#### Fix data  
```{r}

# df_ife %>% names() %>% dput()

vars <- c("d13CVPDB", "d15NAIR", "W.C", 
          "W.N", "C.N.W.Ratio", "Trophic.level", "Average.SD.ARU")

for (var in vars){
  df_ife[[var]] <- as.numeric(gsub(",", ".", df_ife[[var]]))
}

```


#### Substances = d13C and d15N    

```{r}

ggplot(df_ife, aes(d13CVPDB, d15NAIR, color = Specification)) +
  geom_point()

```



## .
## .
## .
## .
