---
title: "100_Labware_Milfersk_2018"
format: html
---

## Read samples from Labware LIMS   
- Script 100_Labware_[program]_[year]     
- function get_labware_samples    
- for comparing samples in Labware with samples in report, and (later) with other data   


## 1. Settings  

### For filenames  
```{r}

# program_year <- "Milfersk_2018"
# current_year <- 2018

```

## 2. Packages and functions  
```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)  
library(knitr)     # for kable
# library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```

## 3. Get all samples  

### Find projects   
```{r}

# "milfersk" mentioned in PROSJEKT    
df_proj1 <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%milfersk%'")

# Samples with "17197" mentioned in PROSJEKT    
df_proj2 <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%17197%'")

df_proj3 <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%210136%'")

df_proj4 <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%13223%'")

proj <- unique(c(df_proj1$PROSJEKT, df_proj2$PROSJEKT, df_proj3$PROSJEKT, df_proj4$PROSJEKT))
oppdrag <- unique(c(df_proj1$ANALYSEOPPDRAG, df_proj2$ANALYSEOPPDRAG, df_proj3$ANALYSEOPPDRAG, df_proj4$ANALYSEOPPDRAG))

proj


```


### Get samples  
```{r}

# Search using ANALYSEOPPDRAG (PROEJCT is problematic due to charcters)
df_lab_all <- get_nivabase_selection(
  "*", "LABWARE_CHECK_SAMPLE", "ANALYSEOPPDRAG", oppdrag, values_are_text = TRUE) %>%
  mutate(Year_id = substr(TEXT_ID, 4,7), Year_sample = year(SAMPLED_DATE))

df_lab_all %>%
  count(PROSJEKT, Year_id)

```

## 3. Tables  

### Stations 

#### Stations all years   
```{r}

df_lab_all %>%
  group_by(AQUAMONITOR_ID, AQUAMONITOR_CODE, AQUAMONITOR_NAME) %>%
  summarize(
    years = paste(sort(unique(Year_id)), collapse = ", "),
    n = n()
  )

```


#### Stations by year  
```{r}

df_lab_all %>%
  group_by(Station = paste(AQUAMONITOR_ID, AQUAMONITOR_CODE, AQUAMONITOR_NAME), Year_id) %>%
  summarize(
    Date_given = paste0(
      sum(!is.na(SAMPLED_DATE)), "/", length(SAMPLED_DATE)),
    .groups = "drop"
  ) %>%
  pivot_wider(names_from = Year_id, values_from = Date_given, names_sort = TRUE)


#  count(AQUAMONITOR_NAME, Year_id = substr(TEXT_ID, 4,7), Date_given = !is.na(SAMPLED_DATE))
  

```

#### Ørret (trout Salmo trutta) stations and tissues   

* Station ID changed between 2018-2020 and 2021    
* No liver samples recorded for 2019 (liver samples were taken acc. report, but for PFAS only, which is sent to NILU)  
* Rogn and Melke should have been taken in 2020 samples (according to report)  
* Rogn and Melke occurs in 2021 - is that correct??   
* No Feminden in 2022 (correct?)  

```{r}

df_lab_all %>%
  filter(grepl("ørret", AQUAMONITOR_NAME, ignore.case = TRUE)) %>% 
  count(Station = paste(AQUAMONITOR_ID, AQUAMONITOR_CODE, AQUAMONITOR_NAME), TISSUE, Year_id) %>%
  arrange(Year_id, Station) %>%
  pivot_wider(names_from = Year_id, values_from = n, names_sort = TRUE)

```


#### Stations  
```{r}

df_lab %>%
  count(AQUAMONITOR_ID, AQUAMONITOR_CODE, AQUAMONITOR_NAME) 

```


#### Benthic food chain     

* None

#### Overvann samples  

* None


## 5. Save  

```{r}

fn <- paste0("Data/100_dat_lab_", program_year, "_", today(tzone = "UTC"), ".rds")
saveRDS(df_lab, fn)

```
