---
title: "100_Labware_Milfersk_2017"
format: html
---

## Read samples from Labware LIMS   
- Script 100_Labware_[program]_[year]     
- function get_labware_samples    
- for comparing samples in Labware with samples in report, and (later) with other data   


## 1. Settings  

### For filenames  
```{r}

program_year <- "Milfersk_2017"
current_year <- 2017

```

## 2. Packages and functions  
```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)  
library(knitr)     # for kable
# library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```

## 3. Get all samples  

### Find projects   
```{r}

# "milfersk" mentioned in PROSJEKT    
df_proj1 <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%milfersk%'")

# Samples with "17197" mentioned in PROSJEKT    
df_proj2 <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%17197%'")

df_proj3 <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where lower(PROSJEKT) like '%210136%'")

proj <- unique(c(df_proj1$PROSJEKT, df_proj2$PROSJEKT, df_proj3$PROSJEKT))
oppdrag <- unique(c(df_proj1$ANALYSEOPPDRAG, df_proj2$ANALYSEOPPDRAG, df_proj3$ANALYSEOPPDRAG))

proj


```


### Get samples, all    
```{r}

# Search using ANALYSEOPPDRAG (PROEJCT is problematic due to charcters)
df_lab_all <- get_nivabase_selection(
  "*", "LABWARE_CHECK_SAMPLE", "ANALYSEOPPDRAG", oppdrag, values_are_text = TRUE) %>%
  mutate(
    Year_id = substr(TEXT_ID, 4,7),
    Year_sample = year(SAMPLED_DATE))

xtabs(~Year_id + Year_sample, df_lab_all) 

df_lab_all

```

## 4. Samples in current year       
```{r}

df_lab <- df_lab_all %>% 
  filter(Year_id == current_year)  

```

### Tables  

#### Stations  
```{r}

df_lab %>%
  count(AQUAMONITOR_ID, AQUAMONITOR_CODE, AQUAMONITOR_NAME) 

```


#### Stations 

```{r}

df_lab %>%
  count(SAMPLE_TYPE, SPECIES, TISSUE) 

``` 
#### Stations trout     

- The samples with 'Ørret lever' and 'Ørret muskel' are NR-2017-12146 - NR-2017-12151
'

```{r}

df_lab %>%
  filter(SPECIES == "Salmo trutta") %>%
  mutate(
    Desc = case_when(
      substr(DESCRIPTION, 3, 3) == "-" ~ substr(DESCRIPTION, 1, 2),
      TRUE ~ DESCRIPTION)
  ) %>%
  count(TISSUE, Desc) 

``` 


## 5. Save  

```{r}

fn <- paste0("Data/100_dat_lab_", program_year, "_", today(tzone = "UTC"), ".rds")
saveRDS(df_lab, fn)

```
