---
title: "100_Labware_Urban_2023"
format: html
---

## Read samples from Labware LIMS   
- Script 1**_Labware_[program]_[year]     
- function get_labware_samples    
- for comparing samples in Labware with samples in report, and (later) with other data   


## 1. Settings  

### For filenames  
```{r}

program_year <- "Urban_2023"
current_year <- 2023

```

## 2. Packages and functions  
```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)  
library(knitr)     # for kable
library(glue)      # alternative to paste()  
# library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()
source("0001_Functions_sql_and_more.R")

knitr::opts_chunk$set(results = 'hold')

```

## 3. Get all samples  

### Find projects/samples     
```{r}

df_proj <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where PROSJEKT = 'O 210135;ANA - Miljøgifter i en urban fjord (2021-2026)'") 

df_proj <- get_nivabase_data("select PROSJEKT, ANALYSEOPPDRAG from NIVADATABASE.LABWARE_CHECK_SAMPLE where PROSJEKT like '%urban%'") %>%
  distinct(PROSJEKT, ANALYSEOPPDRAG)

df_samples_all <- get_nivabase_data("select * from NIVADATABASE.LABWARE_CHECK_SAMPLE where PROSJEKT like '%urban%'")

df_samples_all <- df_samples_all %>%
  mutate(
    year = year(SAMPLED_DATE),
    year_from_textid = substr(df_samples_all$TEXT_ID, 4, 7) %>% as.numeric())

ggplot(df_samples_all, aes(x = SAMPLED_DATE, fill = factor(year_from_textid))) +
  geom_histogram()
# Seal blubber samples in "early 2021" which has TEXT_ID for 2022 must have wrong SAMPLE_DATE   
# See below

# xtabs(~addNA(year) + addNA(year_from_textid), df_samples_all)  
# NOTE: 2022 data (acc. to text ID) with no date

```


### Get all samples 

- ANALYSEOPPDRAG set by comparing with excel 'Samlede resultater 2022_Urbanfjord'   
- https://niva365.sharepoint.com/:x:/r/sites/Miljgifterienurbanfjord2021-2026/_layouts/15/Doc.aspx?sourcedoc=%7B3C315A97-1A5F-4B0C-BDA5-C87DD05BF0A7%7D&file=Samlede%20resultater%202022_Urbanfjord.xlsx&action=default&mobileredirect=true   




## 4. Samples in current year       
```{r}

df_lab <- df_samples_all %>% 
  dplyr::filter(year_from_textid %in% current_year) %>% # nrow()
  arrange(TEXT_ID)

```

### Check columns that should be there     

```{r}

check <- check_lims_samples(df_lab)

```

### Tables  

#### Stations  

* NOTE: Station lacking for 79 samples ! 

```{r}

df_lab %>%
  count(AQUAMONITOR_ID, AQUAMONITOR_CODE, AQUAMONITOR_NAME) 

```

#### Samples where station is lacking  
#### Samples where station is lacking and species not given   

* None lacking except 'blankprøve'  
* See 2022 for examples of code  


#### Check description for freshwater samples   

* all are blanks  

```{r}

df_lab %>%
  filter(is.na(AQUAMONITOR_ID) & SAMPLE_TYPE == "FERSKVANN") %>% # View()
  count(TEXT_ID, DESCRIPTION) 

```

#### Check dates  

* OK 

```{r}

df_lab %>%
  count(AQUAMONITOR_ID, AQUAMONITOR_CODE, AQUAMONITOR_NAME, SAMPLED_DATE)

```

## 5. Water samples with water and particular phase  

### Table  

```{r}

df_lab %>%
  filter(
    grepl("(V)", DESCRIPTION, fixed = TRUE) |
    grepl("(W)", DESCRIPTION, fixed = TRUE) |
    grepl("(P)", DESCRIPTION, fixed = TRUE),
  ) %>%
  filter(
    !grepl("Blank", DESCRIPTION, fixed = TRUE)
  ) %>%
  distinct(TEXT_ID, AQUAMONITOR_CODE, DESCRIPTION, SAMPLED_DATE)

```

### Add Nivabase sample ID for water    

```{r}

lookup_nbid_water <- get_nivabase_selection(
  c("WATER_SAMPLE_ID, LABWARE_SAMPLE_NUMBER, LABWARE_TEXT_ID"),
  "LABWARE_WSID",
  "LABWARE_TEXT_ID", unique(df_lab$TEXT_ID), values_are_text = TRUE)

df_lab <- df_lab %>%
  left_join(lookup_nbid_water,
            by = c("TEXT_ID" = "LABWARE_TEXT_ID"), relationship = "many-to-one")

```

### Check WATER_SAMPLE_ID  

* same WATER_SAMPLE_ID for both phases  

```{r}

df_lab %>%
  filter(SAMPLE_TYPE == "FERSKVANN") %>%
  select(TEXT_ID, DESCRIPTION, WATER_SAMPLE_ID) 

```


## 6. Save  

```{r}

fn <- glue::glue("Data/100_dat_lab_{program_year}_{today(tzone = 'UTC')}.rds")
fn

saveRDS(df_lab, fn)


if (FALSE){
  
  # Read back
  pattern <- glue::glue("100_dat_lab_{program_year}.*rds")
  fns <- dir("Data", pattern) %>% rev()
  fns
  fn <- paste0("Data/", head(fns, 1))
  df_lab <- readRDS(fn)
  table(df_lab$ANALYSEOPPDRAG)
  
  # Get ANALYSEOPPDRAG for SQL
  options(useFancyQuotes = FALSE)
  df_lab$ANALYSEOPPDRAG %>% unique() %>% sQuote() %>% paste(collapse = ", ")
  
}
```

## APPENDIX 1: Labware values   

### One water sample with water + particle phase  

* Particle phase are always AQUAMONITOR_UNIT = NA   
* Correct: NG_P_L_2 -> ng/L 
* Correct: NG_P_ML -> ug/L 
* But some have UNITS = "NG_P_L_2" but AQUAMONITOR_UNIT "µg/L" or "µg/l"  

```{r}

check <- get_nivabase_selection(
  "*", "LABWARE_IMPORT", "TEXT_ID", c("NR-2023-06748", "NR-2023-06749"), values_are_text = TRUE)   

check %>%
  count(DESCRIPTION, UNITS, AQUAMONITOR_UNIT)

```

### Check up on data with discprepancy between UNITS and AQUAMONITOR_UNIT

```{r}

check %>%
  filter(UNITS == "NG_P_L_2" &  AQUAMONITOR_UNIT %in% c("µg/L", "µg/l")) %>%
  select(DESCRIPTION, REPORTED_NAME, NUMERIC_ENTRY, UNITS, AQUAMONITOR_UNIT, VERDI_I_AQUAMONITOR)

```

## APPENDIX 2: SQL examples  

```{r}

# all water samples in 2024, all columns  
df_samples_water_all <- get_nivabase_data("
select *
from NIVADATABASE.LABWARE_CHECK_SAMPLE 
where PROSJEKT like '%urban%' 
and substr(TEXT_ID,4,4) = '2023'
and SAMPLE_TYPE = 'FERSKVANN'
")

# LIMS sample data
# all accepted water samples with station in 2024, selected columns  
df_samples_water <- get_nivabase_data("
select a.TEXT_ID, a.DESCRIPTION, a.AQUAMONITOR_CODE, b.WATER_SAMPLE_ID 
from NIVADATABASE.LABWARE_CHECK_SAMPLE a
left join NIVADATABASE.LABWARE_WSID b on TEXT_ID = LABWARE_TEXT_ID 
where PROSJEKT like '%urban%' 
and substr(TEXT_ID,4,4) = '2023'
and SAMPLE_TYPE = 'FERSKVANN'
and AQUAMONITOR_CODE is not null
and STATUS = 'Ferdig'
order by TEXT_ID
")
write_excel_csv(df_samples_water, "Data/100_dat_lab_Urban_2023_2024-30-05.csv")


# LIMS sample data from one sample with both phases 
check <- get_nivabase_data("
select a.TEXT_ID, a.DESCRIPTION, a.AQUAMONITOR_CODE, a.SAMPLED_DATE, b.WATER_SAMPLE_ID 
from NIVADATABASE.LABWARE_CHECK_SAMPLE a
left join NIVADATABASE.LABWARE_WSID b on TEXT_ID = LABWARE_TEXT_ID 
where TEXT_ID in ('NR-2023-06748', 'NR-2023-06749') 
order by TEXT_ID
")

# LIMS concentration data from one sample with both phases 
check <- get_nivabase_data("
  select *  
  from NIVADATABASE.LABWARE_IMPORT 
  where TEXT_ID in ('NR-2023-06748', 'NR-2023-06749') 
  ")

# Summarize units for one sample with both phases  
get_nivabase_data("
select TEXT_ID, DESCRIPTION, UNITS, AQUAMONITOR_UNIT, count(*)
from NIVADATABASE.LABWARE_IMPORT 
where TEXT_ID in ('NR-2023-06748', 'NR-2023-06749')
group by TEXT_ID, DESCRIPTION, UNITS, AQUAMONITOR_UNIT
order by TEXT_ID, DESCRIPTION, UNITS, AQUAMONITOR_UNIT; 
")

# wrong units linked?
get_nivabase_data("
select DESCRIPTION, REPORTED_NAME, NUMERIC_ENTRY, UNITS, AQUAMONITOR_UNIT
from NIVADATABASE.LABWARE_IMPORT 
where TEXT_ID in ('NR-2023-06748', 'NR-2023-06749')
and UNITS = 'NG_P_L_2' 
and AQUAMONITOR_UNIT not in ('ng/L', 'ng/l');
")

# lacking unit
get_nivabase_data("
select DESCRIPTION, REPORTED_NAME, NUMERIC_ENTRY, UNITS, AQUAMONITOR_UNIT
from NIVADATABASE.LABWARE_IMPORT 
where TEXT_ID in ('NR-2023-06748', 'NR-2023-06749')
and UNITS = 'NG_P_L_2' 
and AQUAMONITOR_UNIT is null;
")

# Nivabase dta for one sample with both phases   
# They all have two methods with same name  
df_nb_example <- get_nivabase_data("
select a.SAMPLE_DATE, a.DEPTH1, a.WATER_SAMPLE_ID, b.LABWARE_TEXT_ID, c.VALUE, c.FLAG1, c.METHOD_ID, d.NAME
from NIVADATABASE.WATER_SAMPLES a
left join NIVADATABASE.LABWARE_WSID b on a.WATER_SAMPLE_ID = b.WATER_SAMPLE_ID 
left join NIVADATABASE.WATER_CHEMISTRY_VALUES c on a.WATER_SAMPLE_ID = c.WATER_SAMPLE_ID 
left join NIVADATABASE.METHOD_DEFINITIONS d on c.METHOD_ID = d.METHOD_ID  
where b.LABWARE_TEXT_ID in ('NR-2023-06748', 'NR-2023-06749');
")

table(df_nb_example$NAME) |> table()
table(df_nb_example$METHOD_ID) |> table()

```
### Check latest updates  
```{r}

check <- get_nivabase_data("
select * from all_tab_modifications where TABLE_OWNER in 'NIVADATABASE' and TIMESTAMP > (current_date - INTERVAL '1' DAY);
")

check2 <- get_nivabase_data("
select * from NIVADATABASE.SEARCH_UPDATE where ENTERED_DATE > (current_date - INTERVAL '1' DAY);
")

```

