---
title: "Script 200_Nivabase_Urban_2024"
format: html
---

3b. Read data from Nivabase   
- Script 200_Nivabase_[program]_[year]     
- functions  
    - get_stations_from_project_id  
    - get_specimens_from_stationdata   
    - get_biota_chemistry  
    - get_water_chemistry_from_stations  
    - get_sediment_chemistry_from_stations  
    
## 1. Setup  

### For filenames  
```{r}

program_year <- "Urban_2024"
sampling_year <- 2024
project_id <- 8551      # see below

```

### Labware data  

```{r}

# dat_lab <- readRDS("Data/100_dat_lab_Milfersk_2022_2023-06-22.rds")

```


### Packages and functions  
```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)
library(knitr)     # for kable
library(glue)      # for glue
# library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```


## 2. Stations (all years)     

```{r}

df_projects <- get_projects()   

# df_stations <- get_stations_from_project("MILFERSK", exact = FALSE)
df_stations <- get_stations_from_project("urban", exact = FALSE)
df_stations <- get_stations_from_project("Urban fjord", exact = TRUE)

nrow(df_stations)

```

## 3. Samples from Labware, script 100      

* NOTE: No water - as stations haven't been defined for the water samples (see script 100)   

```{r}

dat_lims_1 <- readRDS("Data/100_dat_lab_Urban_2024_2025-06-15.rds") # HARD-CODED

# table(dat_lims_1$SAMPLE_TYPE)

lookup_id_biota <- get_nivabase_selection(
  "LABWARE_TEXT_ID, BIOTA_SAMPLE_ID",
  "LABWARE_BSID",
  "LABWARE_TEXT_ID",
  subset(dat_lims_1, SAMPLE_TYPE == "BIOTA")$TEXT_ID, 
  values_are_text = TRUE
)

lookup_id_water <- get_nivabase_selection(
  "LABWARE_TEXT_ID, WATER_SAMPLE_ID",
  "LABWARE_WSID",
  "LABWARE_TEXT_ID",
  subset(dat_lims_1, SAMPLE_TYPE == "FERSKVANN")$TEXT_ID, 
  values_are_text = TRUE
)

# Add NIVAbasen IDs
dat_lims <- dat_lims_1 %>% 
  left_join(lookup_id_biota, by = c("TEXT_ID" = "LABWARE_TEXT_ID"), 
            relationship = "one-to-one") %>% 
  left_join(lookup_id_water, by = c("TEXT_ID" = "LABWARE_TEXT_ID"),
            relationship = "one-to-one")

```

## 4. Water data  

### Samples  

```{r}

df_samples_water <- get_nivabase_selection(
  "*",
  "WATER_SAMPLES",
  "WATER_SAMPLE_ID",
  subset(dat_lims, !is.na(WATER_SAMPLE_ID))$WATER_SAMPLE_ID
)
message("No. of samples, water: ", nrow(df_samples_water))


```

### Values  

```{r}

#
## water data  
#

df_values_water_1 <- get_nivabase_selection(
  "WATER_SAMPLE_ID, METHOD_ID, VALUE, FLAG1, VALUE_ID",
  "WATER_CHEMISTRY_VALUES",
  "WATER_SAMPLE_ID",
  subset(dat_lims, !is.na(WATER_SAMPLE_ID))$WATER_SAMPLE_ID
)
message("No. of measurements, water: ", nrow(df_values_water_1))

df_methods_water <- get_nivabase_selection(
  "METHOD_ID, NAME, UNIT, LABORATORY, METHOD_REF, MATRIX_ID",
  "METHOD_DEFINITIONS",
  "METHOD_ID",
  unique(df_values_water_1$METHOD_ID)
)

df_values_water_2 <- df_values_water_1 %>% 
  left_join(df_methods_water, by = join_by(METHOD_ID))

# 'dat_lims' have several rows with the same WATER_SAMPLE_ID
# dat_lims %>% 
#   add_count(WATER_SAMPLE_ID) %>% 
#   filter(n > 1) %>% 
#   arrange(WATER_SAMPLE_ID)

# Must summarize
dat_lims_summ <- dat_lims %>% 
  group_by(WATER_SAMPLE_ID) %>% 
  summarise(
    TEXT_ID = paste(TEXT_ID, collapse = ", "),
    DESCRIPTION = paste(DESCRIPTION, collapse = " | ")
  )
            
df_values_water <- df_values_water_2 %>% 
  left_join(dat_lims_summ, by = join_by(WATER_SAMPLE_ID))

```



#### Check water samples 1

* Check that water samples are correctly entered  
* Check one station, one parameter  

```{r}

# Sample IDs for one station
id <- dat_lims %>% 
  filter(AQUAMONITOR_CODE == "Urban #2") %>% 
  pull(WATER_SAMPLE_ID)

# Values for this station, one parameter
df_values_water_2 %>% 
  filter(WATER_SAMPLE_ID %in% id & NAME %in% c("PFOA", "PFOA-P")) 

```

#### Check water samples 2

* Check one station, one parameter  

```{r}

# Sample IDs for one station
id <- dat_lims %>% 
  filter(AQUAMONITOR_CODE == "Urban #2") %>% 
  pull(WATER_SAMPLE_ID)

# 13793769 13793769

# Values for this station, one parameter
dat_temp <- df_values_water_2 %>% 
  filter(WATER_SAMPLE_ID %in% id) %>% 
  mutate(
    NAME_orig = NAME,
    NAME = sub("-P", "", NAME_orig, fixed = TRUE),
    Particle_phase = grepl("-P", NAME_orig)
  )
tab <- xtabs(~NAME + Particle_phase, dat_temp)
# tab[1:10,]

check <- apply(tab, 1, sum)
table(check)
# all values should be 2

```
#### Check water samples 3

* Check all stations, all parameters   
* Most stations (26)have only water phase of the following:  
    - Diklormetan, Klorheksidin, Triklormetan (kloroform)  
* One station have only one phase for a few spubstances             

```{r}

# Full data set
dat_temp <- df_values_water_2 %>% 
  mutate(
    NAME_orig = NAME,
    NAME = sub("-P", "", NAME_orig, fixed = TRUE),
    Particle_phase = grepl("-P", NAME_orig)
  )

tab <- xtabs(~NAME + Particle_phase + WATER_SAMPLE_ID, dat_temp)

# test
# str(tab)
# tab[1:3, , 1:3]
# tab[1:5, , "13793769"]

check1 <- apply(tab, c(1,3), sum)
# check1[1:3, 1:3]
table(check1)
# all values should be 0 or 2

# count those with 1
check2 <- check1 == 1
if (sum(check2)){
  warning(sum(check2), " samples x parameters lack particle or water phase")
}

# Check stations lacking particle or water phase for at least on substance
check3 <- apply(check1==1, 2, sum)

# Check parameters for one station
id <- "13793743"
# tab[,,id]
pars_with_one_phase <- apply(tab[,,id], 1, sum) == 1
# tab[pars_with_one_phase, , id]

# Check parameters lacking particle or water phase for at least on substance
check4 <- apply(check1==1, 1, sum)
check4[check4 > 0]

```

#### Klorheksidin  

* Klorheksidin (Chlorhexidine) in filtered + unfiltered water, but also in unfiltered water for source tracing   
* Diklormetan (Dichloromethane) and "Triklormetan (kloroform)" (trichloromethan): only unfiltered water for source tracing
```{r}

# Klorheksidin
dat_temp2 <- dat_temp %>% 
  filter(NAME == "Klorheksidin") %>% 
  count(WATER_SAMPLE_ID, Particle_phase) 

xtabs(~WATER_SAMPLE_ID + Particle_phase, dat_temp2)

# Diklormetan
dat_temp2 <- dat_temp %>% 
  filter(NAME == "Diklormetan") %>% 
  count(WATER_SAMPLE_ID, Particle_phase) 

xtabs(~WATER_SAMPLE_ID + Particle_phase, dat_temp2)

# Triklormetan (kloroform)
dat_temp2 <- dat_temp %>% 
  filter(NAME == "Triklormetan (kloroform)") %>% 
  count(WATER_SAMPLE_ID, Particle_phase) 

xtabs(~WATER_SAMPLE_ID + Particle_phase, dat_temp2)

```

## 5. Biota data  

### Samples  

```{r}

df_samples_biota_1 <- get_nivabase_selection(
  "SAMPLE_ID, TISSUE_ID, TAXONOMY_CODE_ID, STATION_ID, SAMPLE_DATE, SAMPLE_NO, REPNO",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  subset(dat_lims, !is.na(BIOTA_SAMPLE_ID))$BIOTA_SAMPLE_ID
)

lookup_tissue <- get_nivabase_data(
  "select TISSUE_ID, TISSUE_NAME from NIVADATABASE.BIOTA_TISSUE_TYPES")

df_samples_biota <- df_samples_biota_1 %>% 
  left_join(lookup_tissue, join_by(TISSUE_ID), relationship = "many-to-one")

mean(is.na(df_samples_biota$STATION_ID))
table(addNA(df_samples_biota$STATION_ID))
mean(is.na(df_samples_biota$SAMPLE_DATE))
table(addNA(df_samples_biota$SAMPLE_DATE))

message("No. of samples, biota: ", nrow(df_samples_biota))

```

### Specimens  

```{r}

df_samples_specimens <- get_nivabase_selection(
  "SAMPLE_ID, SPECIMEN_ID",
  "BIOTA_SAMPLES_SPECIMENS",
  "SAMPLE_ID",
  subset(dat_lims, !is.na(BIOTA_SAMPLE_ID))$BIOTA_SAMPLE_ID
)

df_specimens <- get_nivabase_selection(
  "STATION_ID, SPECIMEN_ID, SPECIMEN_NO, DATE_CAUGHT, TAXONOMY_CODE_ID",
  "BIOTA_SINGLE_SPECIMENS",
  "SPECIMEN_ID",
  subset(df_samples_specimens, !is.na(SPECIMEN_ID))$SPECIMEN_ID
) %>% 
  rename(STATION_ID_specimen = STATION_ID)

df_species_1 <- get_nivabase_selection(
  "TAXONOMY_CODE_ID, CODE, NIVA_TAXON_ID",
  "TAXONOMY_CODES",
  "TAXONOMY_CODE_ID",
  subset(df_specimens, !is.na(TAXONOMY_CODE_ID))$TAXONOMY_CODE_ID
)

df_species_2 <- get_nivabase_selection(
  "NIVA_TAXON_ID, LATIN_NAME",
  "TAXONOMY",
  "NIVA_TAXON_ID",
  subset(df_species_1, !is.na(NIVA_TAXON_ID))$NIVA_TAXON_ID
)

df_specimens_all <- df_samples_specimens %>% 
  full_join(df_specimens, join_by(SPECIMEN_ID)) %>% 
  left_join(df_species_1, join_by(TAXONOMY_CODE_ID), relationship = "many-to-one") %>% 
  left_join(df_species_2, join_by(NIVA_TAXON_ID), relationship = "many-to-one") 

df_specimens <- df_specimens_all %>% 
  group_by(SAMPLE_ID) %>% 
  summarise(
    STATION_ID_specimen = paste(unique(STATION_ID_specimen), collapse = " | "),
    SPECIMEN_ID = paste(unique(SPECIMEN_ID), collapse = " | "),
    SPECIMEN_NO = paste(unique(SPECIMEN_NO), collapse = " | "),
    DATE_CAUGHT = paste(unique(DATE_CAUGHT), collapse = " | "),
    TAXONOMY_CODE_ID = paste(unique(TAXONOMY_CODE_ID), collapse = " | "),
    CODE = paste(unique(CODE), collapse = " | "),
    LATIN_NAME = paste(unique(LATIN_NAME), collapse = " | "),
  ) %>% 
  arrange(
    STATION_ID_specimen, as.numeric(SPECIMEN_NO))

# message("No. of speci, biota: ", nrow(df_samples_biota))

```


### Values  

```{r}

#
## biota  
#

df_values_biota_1 <- get_nivabase_selection(
  "SAMPLE_ID, METHOD_ID, VALUE, FLAG1, VALUE_ID",
  "BIOTA_CHEMISTRY_VALUES",
  "SAMPLE_ID",
  unique(subset(dat_lims, !is.na(BIOTA_SAMPLE_ID))$BIOTA_SAMPLE_ID)
)
message("No. of measurements, biota: ", nrow(df_values_biota_1))

df_methods_biota <- get_nivabase_selection(
  "METHOD_ID, NAME, UNIT, LABORATORY, METHOD_REF, MATRIX_ID",
  "METHOD_DEFINITIONS",
  "METHOD_ID",
  unique(df_values_biota_1$METHOD_ID)
)

df_values_biota_2 <- df_values_biota_1 %>% 
  left_join(df_methods_biota, by = join_by(METHOD_ID))

# 'dat_lims' have several rows with the same WATER_SAMPLE_ID
# dat_lims %>% 
#   add_count(WATER_SAMPLE_ID) %>% 
#   filter(n > 1) %>% 
#   arrange(WATER_SAMPLE_ID)

# Must summarize (just to be on the safe side, in the case of biota)
dat_lims_summ <- dat_lims %>% 
  group_by(BIOTA_SAMPLE_ID) %>% 
  summarise(
    TEXT_ID = paste(TEXT_ID, collapse = ", "),
    DESCRIPTION = paste(DESCRIPTION, collapse = " | ")
  )
            
df_values_biota <- df_values_biota_2 %>% 
  left_join(dat_lims_summ, by = join_by("SAMPLE_ID" == "BIOTA_SAMPLE_ID"))

```

## 6. Stations  

* Directly from the samples table (not via BIOTA_SINGLE_SPECIMENS)  

```{r}

station_ids <- unique(
  c(
    subset(df_samples_water, !is.na(STATION_ID))$STATION_ID,
    subset(df_samples_biota, !is.na(STATION_ID))$STATION_ID
  )
)

df_stations_1 <-  get_nivabase_selection(
  "STATION_ID, STATION_CODE, STATION_NAME, PROJECT_ID",
  "PROJECTS_STATIONS",
  "STATION_ID",
  station_ids
)
message("No. of station x projects: ", nrow(df_stations_1))

df_stations_proj <-  get_nivabase_selection(
  "PROJECT_ID, PROJECT_NAME",
  "PROJECTS",
  "PROJECT_ID",
  unique(subset(df_stations_1, !is.na(PROJECT_ID))$PROJECT_ID)
)

df_stations_2 <- df_stations_1 %>% 
  left_join(df_stations_proj, by = join_by(PROJECT_ID))

# Could pick projects manually from this
# View(df_stations_2)

# We summarise instead
df_stations_3 <- df_stations_2 %>% 
  group_by(STATION_ID) %>% 
  summarise(
    STATION_CODE = paste(unique(STATION_CODE), collapse = " | "),
    STATION_NAME = paste(unique(STATION_NAME), collapse = " | "),
    PROJECT_ID = paste(unique(PROJECT_ID), collapse = " | "),
    PROJECT_NAME = paste(unique(PROJECT_NAME), collapse = " | ")
  )
message("No. of stations: ", nrow(df_stations_3))
 
df_stations_coord_1 <-  get_nivabase_selection(
  "STATION_ID, GEOM_REF_ID",
  "STATIONS",
  "STATION_ID",
  unique(subset(df_stations_3, !is.na(STATION_ID))$STATION_ID)
)

df_stations_coord_2 <-  get_nivabase_selection(
  "SAMPLE_POINT_ID, LONGITUDE, LATITUDE, EPSG_SYS, X, Y",
  "SAMPLE_POINTS",
  "SAMPLE_POINT_ID",
  unique(subset(df_stations_coord_1, !is.na(GEOM_REF_ID))$GEOM_REF_ID)
)

df_stations <- df_stations_3 %>% 
  left_join(
    df_stations_coord_1, by = "STATION_ID") %>% 
  left_join(
    df_stations_coord_2, by = c("GEOM_REF_ID"="SAMPLE_POINT_ID"))
 
```



## 7. Final datasets  

```{r}

df_water <- df_samples_water %>% 
  left_join(df_stations %>% select(STATION_ID, STATION_CODE, STATION_NAME, LONGITUDE, LATITUDE), 
            by = join_by(STATION_ID), 
            relationship = "many-to-one") %>% 
  left_join(df_values_water, 
            by = join_by(WATER_SAMPLE_ID), 
            relationship = "one-to-many")  

df_biota <- df_samples_biota %>% 
  left_join(df_stations %>% select(STATION_ID, STATION_CODE, STATION_NAME, LONGITUDE, LATITUDE), 
            by = join_by(STATION_ID), 
            relationship = "many-to-one") %>% 
  left_join(df_specimens %>% select(SAMPLE_ID, STATION_ID_specimen, SPECIMEN_ID, LATIN_NAME), 
            by = join_by(SAMPLE_ID), 
            relationship = "many-to-one") %>% 
  left_join(df_values_biota, 
            by = join_by(SAMPLE_ID), 
            relationship = "one-to-many")  

```

## 8. Save  

```{r}

fn_water <- glue::glue("Data/200_dat_water_{program_year}_{today(tzone = 'UTC')}.rds")
saveRDS(df_water, fn_water)
message("Saved water data as ", sQuote(fn_water))

fn_biota <- glue::glue("Data/200_dat_biota_{program_year}_{today(tzone = 'UTC')}.rds")
saveRDS(df_biota, fn_biota)
message("Saved biota data as ", sQuote(fn_biota))

fn_stations <- glue::glue("Data/200_dat_stations_{program_year}_{today(tzone = 'UTC')}.rds")
saveRDS(df_stations, fn_stations)
message("Saved station data as ", sQuote(fn_stations))

if (FALSE){
  
  # Read back
  pattern <- glue::glue("200_dat_water_{program_year}.*rds")
  fns <- dir("Data", pattern) %>% rev()
  fns
  fn <- paste0("Data/", head(fns, 1))
  dat <- readRDS(fn)

}
```


## 9. Known issues  

### a. LATIN_NAME = 'NA'  (literally 'NA')  

* These are mostly insects, but also oligochaetes (earthworms and the like), see table 2 in report  
* Diptera, Ephemeroptera, Trichoptera are insects   
* Vannmiljø:
    - "LatinskNavnID": 89,
    - "VitenskapligNavn": "Insecta (Klasse)",
    - "KategoriNavn": "Klasse",

```{r}

table(addNA(df_specimens$LATIN_NAME))

```


### b. TISSUE_NAME = 'Annet'  

* These are moss samples  
* Should have been given as PLV (plantevev) in LIMS, see script 100  
* Should have been given as TISSUE_ID = 104, TISSUE_NAME = Plantevev in Nivabase  
* Vannmiljø:
    - "MediumID": "BPV",
    - "Name": "Biota plantevev",
    - "Description": "Vevsprøver fra terrestriske planter f.eks. etasjemose og barnåler.  Analyseresultat angis i våtvekt."

```{r}

table(addNA(df_samples_biota$TISSUE_NAME))

```
### c. UNIT = 'NONE'  

* Metals in particle phase have gotten UNIT = 'NONE'  
* NEED TO CHECK WHY THIS IS WRONG IN VANNMILJØ  

```{r}

table(addNA(df_water_1$UNIT))

test1 <- df_water_1 %>% 
  filter(UNIT == "NONE")
nrow(test1)  # 216
xtabs(~NAME + STATION_NAME, test1)

paramnames <- sub("-P$", "", unique(test1$NAME))
test2 <- df_water_1 %>% 
  filter(NAME %in% paramnames)
nrow(test2)  # also 216, so these are correspondiong water phase

# check that it's similar to above
# xtabs(~NAME + STATION_NAME, test2)

xtabs(~NAME + UNIT, test2)

```

## 9. Add Vannmiljø codes for water   

### Prøvetakmetode  

* Klorheksidin (Chlorhexidine) also in unfiltered water for source tracing   

```{r}

lookup_vm_samplingmethod <- fromJSON("https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSamplingMethodList?filter=")

df_water_2 <- df_water_1 %>%
  mutate(
    NAME_nivabase = NAME,
    NAME = sub("-P", "", NAME_nivabase, fixed = TRUE),
    Particle_phase = grepl("-P", NAME_nivabase)
  ) %>% 
  # Also need to mark both particle and water fraction with 'Sample_filtered' 
  # since this needs an extra 
  group_by(
    WATER_SAMPLE_ID, NAME) %>% 
  mutate(
    Particle_phase_n = sum(Particle_phase)) %>% 
  ungroup() %>% 
  mutate(
    Sample_filtered = Particle_phase_n > 0) 
  
df_water_2 %>% filter(NAME == "PFOSA") %>% View()

```


## 10. Add Vannmiljø codes for biota 

### Species  
* All  
    - https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSpeciesList?filter
* Example  
    - https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSpeciesList?filter=Insecta
* LATIN_NAME = 'NA' hardcoded to 89 (Insecta)  
    
```{r}

library(jsonlite)

# test
# fromJSON("https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSpeciesList?filter=Insecta")

lookup_vm_species <- fromJSON("https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSpeciesList?filter=")

df_biota_2 <- df_biota_1 %>% 
  left_join(
    lookup_vm_species %>% select(VitenskapligNavn, LatinskNavnID), 
    by = join_by(LATIN_NAME == VitenskapligNavn), 
    relationship = "many-to-one")

table(addNA(df_biota_2$LatinskNavnID))

sel <- is.na(df_biota_2$LatinskNavnID)
sum(sel)
# table(addNA(df_biota_2$LATIN_NAME[sel]))

if (sum(sel) > 0){
  warning("Species hard-coded where not found")
}

# HARD-CODED - see 8a above 
df_biota_2$LatinskNavnID[sel] <- 89            # HARD-CODED FOR 2024 specifically
df_biota_2$LATIN_NAME[sel] <- "Insecta"        # HARD-CODED FOR 2024 specifically

# Check
table(addNA(df_biota_2$LATIN_NAME))

```


### Tissue    

```{r}

table(df_biota_2$TISSUE_NAME)

sel <- df_biota_2$TISSUE_NAME %in% "Annet"

if (sum(sel) > 0){
  warning("Tissue hard-coded where not found")
  df_biota_2$TISSUE_ID[sel] <- 104                   # HARD-CODED FOR 2024 specifically
  df_biota_2$TISSUE_NAME[sel] <- "Plantevev"         # HARD-CODED FOR 2024 specifically  
  
}

df_convmedium <- get_nivabase_data(
  "select TISSUE_ID, MEDIUM_ID, DATATYPE_ID, STATION_TYPE_ID from VANNMILJO.CONV_MEDIUM")

df_biota_3 <- df_biota_2 %>% 
  left_join(
    df_convmedium %>% select(TISSUE_ID, MEDIUM_ID), 
    by = join_by(TISSUE_ID), 
    relationship = "many-to-one")

sel <- is.na(df_biota_3$MEDIUM_ID)
sum(sel)

if (sum(sel) > 0){
  stop("VN tissues lacking")
} else 
{
  message("All VM tissues OK")
}


```

## APPENDIX: check VANNMILJO tables    


```{r}

get_nivabase_data(
"select OWNER,TABLE_NAME from ALL_TAB_COLUMNS 
where column_name = 'SAMPLE_NUMBER';"
)

get_nivabase_data(
"select OWNER,TABLE_NAME from ALL_TAB_COLUMNS 
where OWNER = 'VANNMILJO';"
)

get_nivabase_data(
"select * from ALL_TAB_COLUMNS 
where OWNER = 'VANNMILJO' and TABLE_NAME = 'CONV_SAMPLING';"
)

```

