---
title: "Script 200_Nivabase_Urban_2024"
format: html
---

3b. Read data from Nivabase   
- Script 200_Nivabase_[program]_[year]     
- functions  
    - get_stations_from_project_id  
    - get_specimens_from_stationdata   
    - get_biota_chemistry  
    - get_water_chemistry_from_stations  
    - get_sediment_chemistry_from_stations  
    
## 1. Setup  

### For filenames  
```{r}

program_year <- "Urban_2024"
sampling_year <- 2024
project_id <- 8551      # see below

```

### Labware data  

```{r}

# dat_lab <- readRDS("Data/100_dat_lab_Milfersk_2022_2023-06-22.rds")

```


### Packages and functions  
```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)
library(knitr)     # for kable
library(glue)      # for glue
# library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```


## 2.  Check existing NIVAbasen data    

```{r}

df_projects <- get_projects()   

# df_stations <- get_stations_from_project("MILFERSK", exact = FALSE)
df_stations <- get_stations_from_project("urban", exact = FALSE)
df_stations <- get_stations_from_project("Urban fjord", exact = TRUE)

nrow(df_stations)

```

### Samples from script 100      

* NOTE: No water - as stations haven't been defined for the water samples (see script 100)   

```{r}

dat_lims <- readRDS("Data/100_dat_lab_Urban_2024_2025-06-15.rds") # HARD-CODED

# table(dat_lims$SAMPLE_TYPE)

lookup_id_biota <- get_nivabase_selection(
  "LABWARE_TEXT_ID, BIOTA_SAMPLE_ID",
  "LABWARE_BSID",
  "LABWARE_TEXT_ID",
  subset(dat_lims, SAMPLE_TYPE == "BIOTA")$TEXT_ID, 
  values_are_text = TRUE
)

lookup_id_water <- get_nivabase_selection(
  "LABWARE_TEXT_ID, WATER_SAMPLE_ID",
  "LABWARE_WSID",
  "LABWARE_TEXT_ID",
  subset(dat_lims, SAMPLE_TYPE == "FERSKVANN")$TEXT_ID, 
  values_are_text = TRUE
)

# Add NIVAbasen IDs
dat_lims_2 <- dat_lims %>% 
  left_join(lookup_id_biota, by = c("TEXT_ID" = "LABWARE_TEXT_ID"), 
            relationship = "one-to-one") %>% 
  left_join(lookup_id_water, by = c("TEXT_ID" = "LABWARE_TEXT_ID"),
            relationship = "one-to-one")

```

### Water data  

#### Samples  


```{r}

df_samples_water <- get_nivabase_selection(
  "*",
  "WATER_SAMPLES",
  "WATER_SAMPLE_ID",
  subset(dat_lims_2, !is.na(WATER_SAMPLE_ID))$WATER_SAMPLE_ID
)
message("No. of samples, water: ", nrow(df_samples_water))


```

%### Values  

```{r}

#
## Water data  
#

df_values_water_1 <- get_nivabase_selection(
  "WATER_SAMPLE_ID, METHOD_ID, VALUE, FLAG1",
  "WATER_CHEMISTRY_VALUES",
  "WATER_SAMPLE_ID",
  subset(dat_lims_2, !is.na(WATER_SAMPLE_ID))$WATER_SAMPLE_ID
)
message("No. of measurements, water: ", nrow(df_values_water_1))

df_methods_water <- get_nivabase_selection(
  "METHOD_ID, NAME, UNIT, LABORATORY, METHOD_REF, MATRIX_ID",
  "METHOD_DEFINITIONS",
  "METHOD_ID",
  unique(df_values_water_1$METHOD_ID)
)

df_values_water_2 <- df_values_water_1 %>% 
  left_join(df_methods_water)

df_values_water_3 <- df_values_water_2 %>% 
  left_join(dat_lims_2)

```

#### Check water samples 1

Check that water samples are correctly entered 1

```{r}

# Sample IDs for one station
id <- dat_lims_2 %>% 
  filter(AQUAMONITOR_CODE == "Urban #2") %>% 
  pull(WATER_SAMPLE_ID)

# Values for this station, one parameter
df_values_water_2 %>% 
  filter(WATER_SAMPLE_ID %in% id & NAME %in% c("PFOA", "PFOA-P")) 

```

#### Check water samples 2

```{r}

# Sample IDs for one station
id <- dat_lims_2 %>% 
  filter(AQUAMONITOR_CODE == "Urban #2") %>% 
  pull(WATER_SAMPLE_ID)

# Values for this station, one parameter
check <- df_values_water_2 %>% 
  filter(WATER_SAMPLE_ID %in% id) %>% 
  mutate(
    NAME_orig = NAME,
    NAME = sub("-P", "", NAME_orig, fixed = TRUE),
    Particle_phase = grepl("-P", NAME_orig)
  )
xtabs(~NAME + Particle_phase, check)

```

#### Get all data    

* NOTE: No water - as stations haven't been defined for the water samples (see script 100)   
```{r}

# Code from script 0010  

# 
## Biota samples
#

df_specimens <- get_specimens_from_stationdata(df_stations, years = sampling_year)
message("No. of specimens, biota: ", nrow(df_specimens))
dat_nb_biota <- get_biota_chemistry(sampling_year, df_specimens, df_stations, report_samples = TRUE)

dat_nb_biota <- dat_nb_biota %>%
  mutate(Year = year(SAMPLE_DATE))

# 1416  

#
## Water samples  
#

df_samples_water <- get_nivabase_selection(
  "*",
  "WATER_SAMPLES",
  "STATION_ID",
  df_stations$STATION_ID, 
  extra_sql = paste("and extraxt(YEAR from SAMPLE_DATE) =", sampling_year)
)
message("No. of samples, water: ", nrow(df_samples_water))

# Adds "SAMPLE_TYPE", "SAMPLE_NUMBER", "DESCRIPTION" and "n_labware" from Labware

# debugonce(get_water_chemistry_from_stations)
dat_nb_water <- get_water_chemistry_from_stations(unique(df_samples_water$STATION_ID), sampling_year,
                                                  project_id = project_id)
message("No. of measurements, water: ", nrow(dat_nb_water))

# 9355  

df_samples_sediment <- get_nivabase_selection(
  "*",
  "SEDIMENT_SAMPLES",
  "STATION_ID",
  df_stations$STATION_ID
)
message("No. of samples, sediment: ", nrow(df_samples_sediment))

dat_nb_sediment <- get_sediment_chemistry_from_stations(df_stations$STATION_ID, years = sampling_year)

if ()
dat_nb_sediment_methods <- get_chemistry_methods(method_id = unique(dat_nb_sediment$METHOD_ID))

dat_nb_sediment <- dat_nb_sediment %>%
  left_join(dat_nb_sediment_methods, by = "METHOD_ID") %>%
  mutate(Year = year(SAMPLE_DATE)) 

# 224

```

### Check biota data  

#### Number of individuals per station  
```{r}

dat_nb_biota %>%
  distinct(LATIN_NAME, STATION_CODE, SAMPLE_NO) %>%
  count(LATIN_NAME, STATION_CODE)

```

#### Number of individuals per station  
```{r}

dat_nb_biota %>%
  distinct(NAME, LATIN_NAME, SAMPLE_NO) %>%
  count(NAME, LATIN_NAME, name = "number_ind") %>%
  pivot_wider(names_from = LATIN_NAME, values_from = number_ind)

```


### Hunt for more (and other) biota data  

* not this year (see 2022 version)  



## 3. Couple Labware TEXT_ID per Nivabase SAMPLE_ID

### Get Labware sample ID     

```{r}

# get_nivabase_data("select * from NIVADATABASE.LABWARE_BSID where rownum < 6")  
# get_nivabase_data("select * from NIVADATABASE.LABWARE_WSID where rownum < 6")  
# get_nivabase_data("select * from NIVADATABASE.LABWARE_SLICE_ID where rownum < 6")  

# unique(dat_nb_biota$SAMPLE_ID)

dat_lab_biota <- get_nivabase_selection(
  c("BIOTA_SAMPLE_ID, LABWARE_SAMPLE_NUMBER, LABWARE_TEXT_ID"),
  "LABWARE_BSID", 
  "BIOTA_SAMPLE_ID", unique(dat_nb_biota$SAMPLE_ID)) 

dat_lab_water <- get_nivabase_selection(
  c("WATER_SAMPLE_ID, LABWARE_SAMPLE_NUMBER, LABWARE_TEXT_ID"),
  "LABWARE_WSID",
  "WATER_SAMPLE_ID", unique(dat_nb_water$WATER_SAMPLE_ID))

dat_lab_sediment <- get_nivabase_selection(
  c("SLICE_ID, LABWARE_SAMPLE_NUMBER, LABWARE_TEXT_ID"),
  "LABWARE_SLICE_ID", 
  "SLICE_ID", unique(dat_nb_sediment$SLICE_ID)) 

```

## 4. Check water data  

### Get labware sample metadata 

#### from sample file

```{r}

fns <- dir("Data", pattern = glue("100_dat_lab_{program_year}.*rds")) %>% rev()
fns
fn <- fns[1]

meta_lab_water_1 <- readRDS(glue("Data/", {fn})) %>%
  filter(TEXT_ID %in% dat_lab_water$LABWARE_TEXT_ID)

nrow(dat_lab_water)
nrow(meta_lab_water_1)

if (nrow(dat_lab_water) > nrow(meta_lab_water_1)){
  
  # Some Nivabase data are not in the 
  dat_lab_water_notfound <- dat_lab_water %>%
    filter(!LABWARE_TEXT_ID %in% meta_lab_water_1$TEXT_ID)
  
  # dat_nb_water %>% 
  #   filter(WATER_SAMPLE_ID %in% dat_lab_water_notfound$WATER_SAMPLE_ID) %>% View()
  dat_nb_water %>% 
    filter(WATER_SAMPLE_ID %in% dat_lab_water_notfound$WATER_SAMPLE_ID) %>% xtabs(~NAME, .)
  # metals, UV, nutrients  
}

```

#### from Labware

```{r}

meta_lab_water <- get_nivabase_selection(
  "*",  
  "LABWARE_CHECK_SAMPLE", 
  "TEXT_ID", 
  dat_lab_water$LABWARE_TEXT_ID, 
  values_are_text = TRUE)

nrow(dat_lab_water)
nrow(meta_lab_water_1)
nrow(meta_lab_water)

```
### Check 1  
```{r}

check_meta <- meta_lab_water %>%
  filter(grepl("Alna River Water 1", DESCRIPTION)) %>%
  select(TEXT_ID, DESCRIPTION)
check_meta  # two lines

check_textid <- check_meta %>%
  pull(TEXT_ID)

check_lookup <- dat_lab_water %>%
  filter(LABWARE_TEXT_ID %in% check_textid)
table(check_lookup$LABWARE_TEXT_ID)   # 2 values
table(check_lookup$WATER_SAMPLE_ID)   # 1 value  

check_id <- check_lookup %>%
  pull(WATER_SAMPLE_ID)
check_id  # two IDs in lookup (BUT THEY ARWE THE SAME.....) 

check_data <- dat_nb_water %>%
  filter(WATER_SAMPLE_ID %in% check_id)

table(check_data$WATER_SAMPLE_ID)
# One ID in the NB data  

```

### Check one 
```{r}

dat_lab_water %>%
  filter(WATER_SAMPLE_ID == 13579823) %>%
  left_join(meta_lab_water,
            by = c("LABWARE_TEXT_ID" = "TEXT_ID"), relationship = "many-to-one") %>%
  select(WATER_SAMPLE_ID, LABWARE_TEXT_ID, DESCRIPTION)

nb_water_one_sample <- dat_nb_water %>% 
  filter(WATER_SAMPLE_ID == 13579823) %>%
  mutate(NAME_type = ifelse(
    stringr::str_sub(NAME, start = -2) == "-P",
    "Particle phase",
    "Water phase")
    )

xtabs(~NAME_type + paste(STATION_CODE, SAMPLE_DATE), nb_water_one_sample)

# View(nb_water_one_sample)

```

### Check one substance

```{r}

nb_water_one_substance <- dat_nb_water %>% filter(grepl("PCB 118", NAME)) 

xtabs(~STATION_CODE + SAMPLE_DATE + NAME, nb_water_one_substance)

# View(nb_water_one_substance)

```


```{r}

table(dat_nb_water$WATER_SAMPLE_ID)
table(dat_nb_water$STATION_CODE)

check <- dat_nb_water %>%
  count(STATION_CODE, WATER_SAMPLE_ID, name = "n_meas")

check %>%
  count(STATION_CODE, name = "n_samples")

```





#### Number of parameters per sample    

* n_samples = 2 for water samples  
    - represents water and particle fractions from the same physical sample    
    - 'Avløpsvann og avløpsslam' and 'Hamar overvann'   
* n_samples = 3 for trout (ørret) and surface water    
    - represents 3 physical samples      

```{r}

dat_nb_biota_names <- dat_nb_biota %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, NAME, Year, name = "n_samples")
df1 <- dat_nb_biota_names %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, n_samples)
# df1

dat_nb_water_names <- dat_nb_water %>%
  mutate(LATIN_NAME = NA, TISSUE_NAME = NA) %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, NAME, Year, name = "n_samples")
df2 <- dat_nb_water_names %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, n_samples)
# df2

dat_nb_sediment_names <- dat_nb_sediment %>%
  mutate(LATIN_NAME = NA, TISSUE_NAME = NA) %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, NAME, Year, name = "n_samples")
df3 <- dat_nb_sediment_names %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, n_samples)
# df3

bind_rows(df1, df2, df3)


```


## 4. Save  

```{r}

saveRDS(dat_nb_biota, paste0("Data/200_dat_nb_biota_", program_year, "_", today(), ".rds"))
saveRDS(dat_nb_water, paste0("Data/200_dat_nb_water_", program_year, "_", today(), ".rds"))
saveRDS(dat_nb_sediment, paste0("Data/200_dat_nb_sediment_", program_year, "_", today(), ".rds"))

if (FALSE){
  files <- dir("Data", "200_dat_nb_sediment_", full.names = TRUE) %>% sort(decreasing = TRUE)
  dat_nb_sediment <- readRDS(files[1])
  table(dat_nb_sediment$SLICE_ID)
}

```



## STUFF NOT USED HERE   


### Add labware TEXT_ID to the data?  

* Not done, as there may be two (or more?) Labware TEXT_ID per Nivabase SAMPLE_ID   

* Demonstrated for biota  

* Duplicate of BIOTA_SAMPLE_ID = 247003  
    - Labware numbers NR-2021-07918 (DESCRIPTION = ØM-L-1) and NR-2021-11104 (DESCRIPTION = ØM-L-BP)  
    - Seems to be two rounds of analyses: Some parameters has a value for DESCRIPTION = ØM-L-1 and value = "T" fpr DESCRIPTION = ØM-L-BP, some have the other  

```{r}

# Expect this to have zero rows:
check <- dat_lab_biota %>% add_count(BIOTA_SAMPLE_ID) %>% filter(n > 1)

if (nrow(check) == 0){
  message("No duplicates of any SAMPLE_ID")
} else {
  stop("Duplicates of any SAMPLE_ID - see script 200_Nivabase_Milfersk_2021")
}

```
