---
title: "Script 200_Nivabase_Milfersk_2022"
format: html
---

3b. Read data from Nivabase   
- Script 200_Nivabase_[program]_[year]     
- functions  
    - get_stations_from_project_id  
    - get_specimens_from_stationdata   
    - get_biota_chemistry  
    - get_water_chemistry_from_stations  
    - get_sediment_chemistry_from_stations  
    
## 1. Setup  

### For filenames  
```{r}

program_year <- "Milfersk_2013-2022"
sampling_year <- 2013:2022

```

### Labware data  

```{r}

# dat_lab <- readRDS("Data/100_dat_lab_Milfersk_2022_2023-06-22.rds")

```


### Packages and functions  
```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)
library(knitr)     # for kable
# library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```


## 2.  Check existing NIVAbasen data    

```{r}

df_projects <- get_projects()   

# df_stations <- get_stations_from_project("MILFERSK", exact = FALSE)
df_stations1 <- get_stations_from_project("MILFERSK", exact = TRUE) %>%
  mutate(Project = "MILFERSK")
df_stations2 <- get_stations_from_project("MILFERSK 2021-2025", exact = TRUE) %>%
  mutate(Project = "MILFERSK 2021-2025")
df_stations3 <- get_stations_from_project("Miljøgifter i store norske innsjøer, 2013. LNR: 667-2014", exact = TRUE) %>%
  mutate(Project = "Store innsjøer")

df_stations <- bind_rows(df_stations1, df_stations2, df_stations3)

nrow(df_stations)

```

#### Get all data    

* NOTE: No water - as stations haven't been defined for the water samples (see script 100)   

```{r}

# Code from script 0010  

# 
## BIOTA  
#

df_specimens <- get_specimens_from_stationdata(df_stations, years = sampling_year)
message("No. of specimens, biota: ", nrow(df_specimens))

dat_nb_biota <- get_biota_chemistry(sampling_year, df_specimens, df_stations, report_samples = TRUE)

df_specimens <- df_specimens %>% mutate(Year = year(DATE_CAUGHT))
dat_nb_biota <- dat_nb_biota %>% mutate(Year = year(SAMPLE_DATE))


# 421  

#
## WATER  
#

df_samples_water <- get_nivabase_selection(
  "*",
  "WATER_SAMPLES",
  "STATION_ID",
  df_stations$STATION_ID
)
message("No. of samples, water: ", nrow(df_samples_water))

# Adds "SAMPLE_TYPE", "SAMPLE_NUMBER", "DESCRIPTION" and "n_labware" from Labware

# debugonce(get_water_chemistry_from_stations)
dat_nb_water <- get_water_chemistry_from_stations(unique(df_samples_water$STATION_ID), sampling_year)

df_samples_water <- df_samples_water %>% mutate(Year = year(SAMPLE_DATE))
dat_nb_water <- dat_nb_water %>% mutate(Year = year(SAMPLE_DATE))

#
## SEDIMENT
#

df_samples_sediment <- get_nivabase_selection(
  "*",
  "SEDIMENT_SAMPLES",
  "STATION_ID",
  df_stations$STATION_ID
)
message("No. of samples, sediment: ", nrow(df_samples_sediment))

dat_nb_sediment <- get_sediment_chemistry_from_stations(df_stations$STATION_ID, years = sampling_year)

dat_nb_sediment_methods <- get_chemistry_methods(method_id = unique(dat_nb_sediment$METHOD_ID))

df_samples_water <- df_samples_water %>% mutate(Year = year(SAMPLE_DATE))

df_samples_sediment <- df_samples_sediment %>% mutate(Year = year(SAMPLE_DATE))
dat_nb_sediment <- dat_nb_sediment %>%
  left_join(dat_nb_sediment_methods, by = "METHOD_ID") %>%
  mutate(Year = year(SAMPLE_DATE)) 


```

### Years  
```{r}

cat("BIOTA\n")
cat("\nSpecimens:\n")
table(df_specimens$Year)
cat("\nMeasurements:\n")
table(dat_nb_biota$Year)

cat("\n===============================\n")
cat("WATER\n")
table(df_samples_water$Year)
table(dat_nb_water$Year)

cat("\n===============================\n")
cat("SEDIMENT\n")
table(df_samples_sediment$Year)
table(dat_nb_sediment$Year)

```

## 2. Tables    


### a. Stations, all years   

* Based on 'df_specimens'  

```{r}

df_specimens %>% 
  left_join(
    df_stations %>% select(STATION_ID, STATION_CODE, STATION_NAME)
    ) %>%
  mutate(Station = paste(STATION_ID, STATION_CODE, STATION_NAME)) %>%
  xtabs(~Station + Year, .)

```

#### a2. Stations, trout   

* Based on 'df_specimens'  

```{r}

df_specimens %>%
  filter(LATIN_NAME == "Salmo trutta") %>%
  left_join(
    df_stations %>% select(STATION_ID, STATION_CODE, STATION_NAME)
    ) %>%
  mutate(Station = paste(STATION_ID, STATION_CODE, STATION_NAME)) %>%
  xtabs(~Station + Year, .)

```

### b. Stations, coordinates    

```{r}

df_geom1 <- get_nivabase_selection(
  "STATION_ID, GEOM_REF_ID",
  "STATIONS",
  "STATION_ID",
  df_stations$STATION_ID
)
df_geom2 <- get_nivabase_selection(
  "SAMPLE_POINT_ID, LONGITUDE, LATITUDE",
  "SAMPLE_POINTS", owner = "NIVA_GEOMETRY",
  "SAMPLE_POINT_ID",
  df_geom1$GEOM_REF_ID
)

if (!"LONGITUDE" %in% names(df_stations))
  df_stations <- df_stations %>%
    left_join(df_geom1, by = "STATION_ID") %>%
    left_join(df_geom2, by = c("GEOM_REF_ID"="SAMPLE_POINT_ID"))
  
ggplot(df_stations, aes(LONGITUDE, LATITUDE, color = Project)) +
  geom_point() +
  facet_wrap(vars(Project))

writexl::write_xlsx(df_stations, "Data/600_Milfersk_stations_nivabasen.xlsx")

```

```{r}

if (FALSE){
  
  library(leaflet)
  
  sel <- df_stations$Project == "Store innsjøer"
  sel <- df_stations$Project == "MILFERSK"
  sel <- df_stations$STATION_CODE %in% c("M_Ø", "M_KL", "M_")
  sel <- grepl("Femunden", df_stations$STATION_NAME)
  
  leaflet() %>%
    addTiles() %>%
    addMarkers(lng = df_stations$LONGITUDE[sel], lat = df_stations$LATITUDE[sel], 
               popup = paste(df_stations$STATION_CODE[sel], "<br>", df_stations$STATION_NAME[sel]))
  
}

```


### c. Look for rows in BIOTA_SAMPLES_SPECIMENS  

```{r}

id_spec <- df_specimens %>%
  filter(Year == 2018) %>%
  pull(SPECIMEN_ID)
cat(length(id_spec), "specimens \n\n")


# Coregonus albula = lagesld
# Osmerus eperlanus = krøkle

cat("===============================\nCheck BIOTA_SAMPLES_SPECIMENS\n===============================\n")

# Alt. 1. By ID 
df_samp_spec <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES_SPECIMENS",
  "SPECIMEN_ID",
  id_spec
)
nrow(df_samp_spec) %>% cat("rows\n")
# zero rows  

# Alt. 2. By date (specimens were added by Roar 08-06-2023)  
df_samp_spec <- get_nivabase_data(paste(
  "select * from NIVADATABASE.BIOTA_SAMPLES_SPECIMENS where ENTERED_BY = 'RBR'",
  "and ENTERED_DATE > to_date('07-06-2023', 'DD-MM-YYYY')")
)
nrow(df_samp_spec) %>% cat("rows\n")
# zero rows  

```


### d. Check from the other side, BIOTA_CHEMISTRY_VALUES     

* As it seems that Roar have imported them   
* Check one example substance, based on entered date  
    - 51 measurements  
    - from excel original data, we see that these are the 45 from fish + 3 'zooplankton' + 3 Mysis  

```{r}

df_meth <- get_nivabase_data("select * from NIVADATABASE.METHOD_DEFINITIONS where NAME = 'PFDoDA'")
# METHOD_ID 33759 

df_val <- get_nivabase_data("select * from NIVADATABASE.BIOTA_CHEMISTRY_VALUES where METHOD_ID = 33759 and ENTERED_DATE > to_date('07-06-2023', 'DD-MM-YYYY')")
cat("df_val:", nrow(df_val), "rows \n")
# 51 rows
# table(df_val$ENTERED_BY)     # Roar
# table(df_val$ENTERED_DATE)   # Added 19. July 2023  

# Get those samples
df_samp <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  unique(df_val$SAMPLE_ID)
)
cat("df_samp:", nrow(df_samp), "rows \n")
# 51 rows

df_samp_spec <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES_SPECIMENS",
  "SAMPLE_ID",
  unique(df_val$SAMPLE_ID)
)
cat("df_samp_spec:", nrow(df_samp_spec), "rows \n")

```

### e. Check how to add Zooplankton epilimnion     

* Coupling BIOTA_SINGLE_SPECIMENS -> TAXONOMY_CODES -> TAXONOMY  
    - 1 to 2: link = TAXONOMY_CODE_ID
    - 2 to 3: link = NIVA_TAXON_ID  

```{r}

#
# "Chironomidae" (used in 2021) has 1 row in TAXONOMY
#
df_taxon_check <- get_nivabase_data(
  "select * from NIVADATABASE.TAXONOMY where LATIN_NAME like '%Chironomid%'")
#
# But 7 rows in TAXONOMY_CODES (different stages etc.)
#
df_taxon_check_code <- get_nivabase_data(
  "select * from NIVADATABASE.TAXONOMY_CODES where NIVA_TAXON_ID = 13537")

#
# Zooplankton epilimnion
#
df_taxon_check_code <- get_nivabase_data(
  "select * from NIVADATABASE.TAXONOMY_CODES where NAME like '%epilimn%'")
# TAXONOMY_CODE_ID = 8941


```



## 3. Couple Labware TEXT_ID per Nivabase SAMPLE_ID

### Get Labware sample ID     

```{r}

# get_nivabase_data("select * from NIVADATABASE.LABWARE_BSID where rownum < 6")  
# get_nivabase_data("select * from NIVADATABASE.LABWARE_WSID where rownum < 6")  
# get_nivabase_data("select * from NIVADATABASE.LABWARE_SLICE_ID where rownum < 6")  

# unique(dat_nb_biota$SAMPLE_ID)

dat_lab_biota <- get_nivabase_selection(
  c("BIOTA_SAMPLE_ID, LABWARE_SAMPLE_NUMBER, LABWARE_TEXT_ID"),
  "LABWARE_BSID", 
  "BIOTA_SAMPLE_ID", unique(dat_nb_biota$SAMPLE_ID)) 
# NO WATER at this stage  
# dat_lab_water <- get_nivabase_selection(
#   c("WATER_SAMPLE_ID, LABWARE_SAMPLE_NUMBER, LABWARE_TEXT_ID"),
#   "LABWARE_WSID", 
#   "WATER_SAMPLE_ID", unique(dat_nb_water$WATER_SAMPLE_ID)) 
dat_lab_sediment <- get_nivabase_selection(
  c("SLICE_ID, LABWARE_SAMPLE_NUMBER, LABWARE_TEXT_ID"),
  "LABWARE_SLICE_ID", 
  "SLICE_ID", unique(dat_nb_sediment$SLICE_ID)) 

```



### Add labware TEXT_ID to the data?  

* Not done, as there may be two (or more?) Labware TEXT_ID per Nivabase SAMPLE_ID   

* Demonstrated for biota  

* Duplicate of BIOTA_SAMPLE_ID = 247003  
    - Labware numbers NR-2021-07918 (DESCRIPTION = ØM-L-1) and NR-2021-11104 (DESCRIPTION = ØM-L-BP)  
    - Seems to be two rounds of analyses: Some parameters has a value for DESCRIPTION = ØM-L-1 and value = "T" fpr DESCRIPTION = ØM-L-BP, some have the other  

```{r}

# Expect this to have zero rows:
check <- dat_lab_biota %>% add_count(BIOTA_SAMPLE_ID) %>% filter(n > 1)

if (nrow(check) == 0){
  message("No duplicates of any SAMPLE_ID")
} else {
  stop("Duplicates of any SAMPLE_ID - see script 200_Nivabase_Milfersk_2021")
}

```

#### Number of parameters per sample    

* n_samples = 2 for water samples  
    - represents water and particle fractions from the same physical sample    
    - 'Avløpsvann og avløpsslam' and 'Hamar overvann'   
* n_samples = 3 for trout (ørret) and surface water    
    - represents 3 physical samples      

```{r}

dat_nb_biota_names <- dat_nb_biota %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, NAME, Year, name = "n_samples")
df1 <- dat_nb_biota_names %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, n_samples)
# df1

dat_nb_water_names <- dat_nb_water %>%
  mutate(LATIN_NAME = NA, TISSUE_NAME = NA) %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, NAME, Year, name = "n_samples")
df2 <- dat_nb_water_names %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, n_samples)
# df2

dat_nb_sediment_names <- dat_nb_sediment %>%
  mutate(LATIN_NAME = NA, TISSUE_NAME = NA) %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, NAME, Year, name = "n_samples")
df3 <- dat_nb_sediment_names %>%
  count(STATION_NAME, LATIN_NAME, TISSUE_NAME, n_samples)
# df3

bind_rows(df1, df2, df3)


```


## 4. Save  

```{r}

saveRDS(dat_nb_biota, paste0("Data/200_dat_nb_biota_", program_year, "_", today(), ".rds"))
saveRDS(dat_nb_water, paste0("Data/200_dat_nb_water_", program_year, "_", today(), ".rds"))
saveRDS(dat_nb_sediment, paste0("Data/200_dat_nb_sediment_", program_year, "_", today(), ".rds"))

if (FALSE){
  files <- dir("Data", "200_dat_nb_sediment_", full.names = TRUE) %>% sort(decreasing = TRUE)
  dat_nb_sediment <- readRDS(files[1])
  table(dat_nb_sediment$SLICE_ID)
}

```


## APPENDIX snippets  
```{r}


df_specimens <- df_specimens %>%
  niRvana::get_s

xtabs(~LATIN_NAME, df_specimens)
xtabs(~Year + LATIN_NAME, df_specimens)

xtabs(~Year + STATION_ID, df_specimens %>% filter(LATIN_NAME == "Salmo trutta"))
xtabs(~TAXONOMY_CODE_ID, df_specimens %>% filter(LATIN_NAME == "Salmo trutta"))

dff <- get_nivabase_selection(
  "*",
  "BIOTA_SINGLE_SPECIMENS",
  "TAXONOMY_CODE_ID" ,
  8840, extra_sql = "AND extract(YEAR from DATE_CAUGHT) IN (2013,2014,2015,2016)"
)

dff <- get_nivabase_data(
  "SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE TAXONOMY_CODE_ID = 8840 AND EXTRACT(YEAR FROM DATE_CAUGHT) IN (2013,2014,2015,2016)"
)

dff <- get_nivabase_data(
  "SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE TAXONOMY_CODE_ID = 8840 AND EXTRACT(YEAR FROM DATE_CAUGHT) = 2013"
)

dff <- get_nivabase_data(
  "SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE TAXONOMY_CODE_ID = 8840 AND EXTRACT(YEAR FROM DATE_CAUGHT) IN (2013,2014,2015,2016)"
)

dff <- get_nivabase_data(
  "SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE STATION_ID in 67415 = 8840 AND EXTRACT(YEAR FROM DATE_CAUGHT) IN (2013,2014,2015,2016)"
)

dff <- get_nivabase_data(
  "SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE TAXONOMY_CODE_ID = 8840 AND STATION_ID in (67414,67415)"
)

dff <- get_nivabase_data(
  "SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE STATION_ID in (67414,67415)"
)



SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE TAXONOMY_CODE_ID = 8840 AND EXTRACT(YEAR FROM DATE_CAUGHT) IN (2013,2014,2015,2016)
SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE TAXONOMY_CODE_ID = 8840 AND EXTRACT(YEAR FROM DATE_CAUGHT) IN (2013,2014,2015,2016)
SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE STATION_ID in 67415 = 8840 AND EXTRACT(YEAR FROM DATE_CAUGHT) IN (2013,2014,2015,2016)
SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE TAXONOMY_CODE_ID = 8840 AND STATION_ID in (67414,67415)
SELECT * FROM NIVADATABASE.BIOTA_SINGLE_SPECIMENS WHERE STATION_ID in (67414,67415)
  


```

