---
title: "Script 200_Nivabase_Urban_2019"
format: html
---

3b. Read data from Nivabase   
- Script 200_Nivabase_[program]_[year]     
- functions  
    - get_stations_from_project_id  
    - get_specimens_from_stationdata   
    - get_biota_chemistry  
    - get_water_chemistry_from_stations  
    - get_sediment_chemistry_from_stations  
    
## 1. Setup  

### For filenames  
```{r}

program_year <- "Urban_2019"
sampling_year <- 2019

```

### Labware data  

```{r}

# dat_lab <- readRDS("Data/100_dat_lab_Milfersk_2019_2023-06-22.rds")

```


### Packages and functions  
```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)
library(knitr)     # for kable
# library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```


## 2.  Check existing NIVAbasen data    

```{r}

df_projects <- get_projects()   

# df_stations <- get_stations_from_project("MILFERSK", exact = FALSE)
df_stations <- get_stations_from_project("urban", exact = FALSE)
df_stations <- get_stations_from_project("Urban fjord", exact = TRUE)

nrow(df_stations)

```

#### Get all data    

* NOTE: No water - as stations haven't been defined for the water samples (see script 100)   

```{r}

# Code from script 0010  

# 
## Biota samples
#

df_specimens <- get_specimens_from_stationdata(df_stations, years = sampling_year)
message("No. of specimens, biota: ", nrow(df_specimens))
dat_nb_biota <- get_biota_chemistry(sampling_year, df_specimens, df_stations, report_samples = TRUE)

dat_nb_biota <- dat_nb_biota %>%
  mutate(Year = year(SAMPLE_DATE))

# 421  

#
## Water samples  
#

df_samples_water <- get_nivabase_selection(
  "*",
  "WATER_SAMPLES",
  "STATION_ID",
  df_stations$STATION_ID
)
message("No. of samples, water: ", nrow(df_samples_water))

# Adds "SAMPLE_TYPE", "SAMPLE_NUMBER", "DESCRIPTION" and "n_labware" from Labware

# debugonce(get_water_chemistry_from_stations)
dat_nb_water <- get_water_chemistry_from_stations(unique(df_samples_water$STATION_ID), sampling_year)
message("No. of measurements, water: ", nrow(dat_nb_water))

df_samples_sediment <- get_nivabase_selection(
  "*",
  "SEDIMENT_SAMPLES",
  "STATION_ID",
  df_stations$STATION_ID
)
message("No. of samples, sediment: ", nrow(df_samples_sediment))

dat_nb_sediment <- get_sediment_chemistry_from_stations(df_stations$STATION_ID, years = sampling_year)

dat_nb_sediment_methods <- get_chemistry_methods(method_id = unique(dat_nb_sediment$METHOD_ID))

dat_nb_sediment <- dat_nb_sediment %>%
  left_join(dat_nb_sediment_methods, by = "METHOD_ID") %>%
  mutate(Year = year(SAMPLE_DATE)) 

```

### Water and sediment stations    
```{r}

table(dat_nb_water$STATION_CODE)
table(dat_nb_sediment$STATION_CODE)

```

### Check for example parameters   

- In the export from Vannmiljø, we found NILU parameters to be lacking for Bekkelaget and Alna  
    - See '601_Check_vmexport_Urbanfjord_2023-06-22.qmd'  
- The same seems to be the case in Nivabasen   
    - see below: PFAS are NIVA parameters, UV are IFE parameters, the rest is NILU   

```{r}

# lookup_param <- read_excel("Input_data/Lookup_files/lookup_substancename_biota.xlsx")

# table(dat_nb_sediment$NAME) %>% names() %>% dput()

param_sel <- c("<63 µm", "BDE-100", "Cd", "Dechlorane 601", 
"HCB", "Hg", "SCCP", "PCB-118", "Perfluorheksansulfonat (PFHxS)", 
"Perfluoroktansulfonamid (PFOSA)", "Perfluoroktansyre (PFOA)",
"TBPH (BEH /TBP)", "UV-328")

dat_nb_sediment %>%
  filter(NAME %in% param_sel) %>%
  xtabs(~NAME + STATION_CODE, .)

dat_nb_water %>%
  filter(NAME %in% param_sel) %>% # View()
  xtabs(~NAME + STATION_CODE, .)

```

### Check dates  
- for one example parameter (a NIVA parameter, present in all)    
- Dates are unique  
```{r}

dat_nb_sediment %>%
  filter(NAME %in% "Perfluoroktansyre (PFOA)") %>%
  xtabs(~SAMPLE_DATE + STATION_CODE, .)

dat_nb_water %>%
  filter(NAME %in% "Perfluoroktansyre (PFOA)") %>%
  xtabs(~SAMPLE_DATE + STATION_CODE, .)

```



### Check example water samples
- Just to make sure only date differs for water samples  
- No particle/water phase this year  
```{r}

check1 <- get_nivabase_selection(
  "*",
  "WATER_SAMPLES",
  "WATER_SAMPLE_ID",
  c(11379134,11379135)
)
check2 <- get_nivabase_selection(
  "*",
  "WATER_CHEMISTRY_VALUES",
  "WATER_SAMPLE_ID",
  c(11379134,11379135), extra_sql = " AND METHOD_ID = 33645"
)
check3a <- get_nivabase_selection(
  "*",
  "LABWARE_WSID",
  "WATER_SAMPLE_ID",
  c(11379134,11379135)
)
check3b <- get_nivabase_selection(
  "*",
  "LABWARE_CHECK_SAMPLE",
  "TEXT_ID",
  check3a$LABWARE_TEXT_ID, values_are_text = TRUE
)

```

