---
title: "Script 200_Nivabasefix_Milfersk_2018"
format: html
---

* Starting from the knowledge gotten from '200_Nivabase_Milfersk_2013-2022.qmd'  
    - Nivabasen has only fish specimens, lacking 3 Mysis relicta and 3 Zooplankton epilimnion   
    - Also, there are no BIOTA_SAMPLES_SPECIMENS records linking the specimens to samples  

## 1. Setup  

### For filenames  
```{r}

program_year <- "Milfersk_2018"
sampling_year <- 2018

```

### Labware data  

* Mysis: NR-2018-10628 - NR-2018-10630  
* Zooplankton: NR-2018-10631 - NR-2018-10633


```{r}

dat_lab <- readRDS("Data/100_dat_lab_Milfersk_2018_2023-06-28.rds")

```


### Packages and functions  
```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)
library(knitr)     # for kable
library(ggeasy)
library(forcats)

# library(fuzzyjoin)   # regex_full_join

source("../Milkys/991_Vannmiljo_snippets_functions.R")
source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()
source("0001_Functions_sql_and_more.R")

knitr::opts_chunk$set(results = 'hold')

```


## 2.  Check existing NIVAbasen data    

### Projects  

```{r}

df_projects <- get_projects()   

df_stations <- get_stations_from_project("MILFERSK", exact = TRUE)

nrow(df_stations)

```

### Specimens      

* Data on specimens   
    - Has fish specimens  
    - Lacking 3 Mysis relicta and 3 Zooplankton epilimnion  
* Also, there are no BIOTA_SAMPLES_SPECIMENS records linking the specimens to samples  
    - We will deal with that later   

```{r}

# Code from script 0010  

# 
## BIOTA  
#

df_specimens <- get_specimens_from_stationdata(df_stations, years = sampling_year)
message("No. of specimens, biota: ", nrow(df_specimens))

df_specimens <- df_specimens %>% mutate(Year = year(DATE_CAUGHT))
nrow(df_specimens)

# debugonce(get_biota_chemistry)
# dat_nb_biota <- get_biota_chemistry(sampling_year, df_specimens, df_stations, report_samples = TRUE)
# FAILS because there are no links to BIOTA_SAMPLES_SPECIMENS  

df_specimens %>%
  left_join(df_stations, by = "STATION_ID") %>%
  mutate(St = paste(STATION_ID, STATION_CODE, STATION_NAME)) %>%
  xtabs(~St + DATE_CAUGHT, .)


#
## WATER  
## SEDIMENT
#
# No such samples this year


```

### Get sample and measurement data  

```{r}

df_meth <- get_nivabase_data("select * from NIVADATABASE.METHOD_DEFINITIONS where NAME = 'PFDoDA'")
# METHOD_ID 33759 

df_val_select <- get_nivabase_data("select * from NIVADATABASE.BIOTA_CHEMISTRY_VALUES where METHOD_ID = 33759 and ENTERED_DATE > to_date('07-06-2023', 'DD-MM-YYYY')")
cat("df_val_select:", nrow(df_val_select), "rows \n")
# 51 rows
# table(df_val$ENTERED_BY)     # Roar
# table(df_val$ENTERED_DATE)   # Added 19. July 2023  

# sample_ids <- unique(df_val_select$SAMPLE_ID)
# dput(sample_ids)
sample_ids <- c(248738, 248739, 248787, 248740, 248741, 248788, 248742, 248743, 
                248744, 248745, 248746, 248747, 248748, 248749, 248750, 248751, 
                248752, 248753, 248754, 248755, 248756, 248757, 248758, 248759, 
                248760, 248761, 248762, 248763, 248764, 248765, 248766, 248767, 
                248768, 248769, 248770, 248771, 248772, 248773, 248774, 248775, 
                248776, 248777, 248778, 248779, 248780, 248781, 248782, 248783, 
                248784, 248785, 248786)

# Get those samples
df_samp <- get_nivabase_selection(
  "*",
  "BIOTA_SAMPLES",
  "SAMPLE_ID",
  sample_ids
)

cat("df_samp:", nrow(df_samp), "rows \n")
# 51 rows

# Get all values  
# 
df_val <- get_nivabase_selection(
  "VALUE_ID, SAMPLE_ID, METHOD_ID, VALUE, FLAG1",                          
  "BIOTA_CHEMISTRY_VALUES",
  "SAMPLE_ID", 
  sample_ids)

# METHOD_ID, NAME, UNIT, LABORATORY, MATRIX, MATRIX_ID
df_meth <- get_nivabase_selection(
  "METHOD_ID, NAME, UNIT, LABORATORY, MATRIX, MATRIX_ID", 
  "METHOD_DEFINITIONS",
  "METHOD_ID", 
  unique(df_val$METHOD_ID))

df_val <- df_val %>%
  left_join(df_meth)

```
## 3. Check im order to couple SAMPLE_ID to SPECIEMN ID  

```{r}

df_val %>%
  filter(NAME %in% c("d13CVPDB", "d15NAIR")) %>%
  select(SAMPLE_ID, NAME, VALUE) %>%
  mutate(VALUE = round(VALUE, 2)) %>%
  pivot_wider(id_cols = SAMPLE_ID, names_from = NAME, values_from = VALUE)

# 248738:248739   # ZM-1-2
# 248740:248741   # Mysis 1-2
# 248742:248751   # Krøkle
# 248752:248761   # Lågåsild
# 248762:248776   # Ørret Mjøsa
# 248777:248786   # Ørret Femunden
# 248787          # ZM-3
# 248788          # Mysis 3

lookup_id <- data.frame(
  ID = c(
    paste0("ZM-", 1:3),
    paste0("MM-", 1:3),
    paste0("KM-M-", 1:10),
    paste0("LM-M-", 1:10),
    paste0("ØM-M-", 1:15),
    paste0("ØF-M-", 1:10)
    ),
  SAMPLE_ID = c(248738:248739, 248787,
                248740:248741, 248788,
                248742:248751,
                248752:248761,
                248762:248776,
                248777:248786)
) %>%
  mutate(ID = fct_inorder(factor(ID)))

df_val2 <- df_val %>%
  left_join(lookup_id)



```
### Check some data in order to compare with excel raw data

* Compare with plot in scr 300  
    - Section 3 in scr 300 - plot of BDE47  

```{r}

# BDE47 = 2,2',4,4'-TetBDE 

# table(df_val2$NAME)

df_val2 %>% 
  filter(NAME == "2,2',4,4'-TetBDE") %>%
  ggplot(aes(ID, VALUE, color = is.na(FLAG1))) +
  geom_point()+ 
  ggeasy::easy_rotate_labels("x", angle = -45)

```

### Check specimens  

* Osmerus eperlanus = krøkle  
* Coregonus albula = lågåsild  (lagesild)  
* Salmo trutta = ørret   
* STATION_ID 71086 = Mjøsa 
LATIN_NAME          71085 71086 71088

```{r}

xtabs(~LATIN_NAME + STATION_ID, df_specimens)

  # 71085 M_Ø Mjøsa ørret                       15
  # 71086 M_KL Mjøsa krøkle og lågåsild         20
  # 71088 F_Ø Femunden ørret                    10

df_specimens %>%
  select(LATIN_NAME, STATION_ID, SPECIMEN_NO, SPECIMEN_ID) %>%
  arrange(SPECIMEN_ID, LATIN_NAME, STATION_ID, SPECIMEN_NO)

# Use outpput from the above to add SPECIMEN_ID    
lookup_id <- data.frame(
  ID = c(
    paste0("ZM-", 1:3),        # zooplankton
    paste0("MM-", 1:3),        # Mysis
    paste0("KM-M-", 1:10),     # Krøkle = Osmerus eperlanus
    paste0("LM-M-", 1:10),     # Lågåsild = Coregonus albula
    paste0("ØM-M-", 1:15),     # Ørret = Salmo trutta, Mjøsa
    paste0("ØF-M-", 1:10)      # Ørret = Salmo trutta, Femunden
    ),
  SAMPLE_ID = c(248738:248739, 248787,
                248740:248741, 248788,
                248742:248751,
                248752:248761,
                248762:248776,
                248777:248786),
  SPECIMEN_ID = c(NA, NA, NA,    # not yet added to BIOTA_SINGLE_SPECIMENS (is done below)
                  NA, NA, NA,    # not yet added to BIOTA_SINGLE_SPECIMENS (is done below)
                  333955:333964,
                  333965:333974,
                  333975:333989,
                  333990:333999)
) %>%
  mutate(ID = fct_inorder(factor(ID)))


```


## 4. Add specimens  

### From Labware

```{r}

dat_lab <- readRDS("Data/100_dat_lab_Milfersk_2018_2023-06-28.rds")

table(dat_lab$SAMPLE_TYPE)
table(dat_lab$AQUAMONITOR_CODE)
table(dat_lab$AQUAMONITOR_NAME)
table(dat_lab$SPECIES)
table(dat_lab$SAMPLED_DATE)

xtabs(~AQUAMONITOR_NAME + SPECIES, dat_lab)  

dat_lab %>%
  filter(AQUAMONITOR_NAME == "Mjøsa zooplankton og Mysis") %>%
  arrange(TEXT_ID) %>%
  select(TEXT_ID, SAMPLED_DATE, DESCRIPTION, AQUAMONITOR_ID, AQUAMONITOR_CODE, SPECIES, TISSUE, BIOTA_SAMPLENO)

dat_lab %>%
  filter(AQUAMONITOR_NAME == "Mjøsa zooplankton og Mysis") %>%
  arrange(TEXT_ID) %>%
  select(TEXT_ID, SAMPLED_DATE, DESCRIPTION, AQUAMONITOR_ID, AQUAMONITOR_CODE, SPECIES, TISSUE, BIOTA_SAMPLENO)

```


### Create data to add 

* From script 100, we see that both zooplankton and Mysis are from STATION_ID 71087	(code 'M_ZM')  

```{r}

#
# Zooplankton epilimnion
#
df_taxon_zooplankton <- get_nivabase_data(
  "select * from NIVADATABASE.TAXONOMY_CODES where NAME like '%epilimn%'")
df_taxon_zooplankton$TAXONOMY_CODE_ID
# TAXONOMY_CODE_ID = 8941

#
# Mysis
#
df_taxon_mysis <- get_nivabase_data(
  "select * from NIVADATABASE.TAXONOMY_CODES where NAME like '%Mysis%'")
unique(df_taxon_mysis$TAXONOMY_CODE_ID)  
# TAXONOMY_CODE_ID = 14089


# make_sql_single_specimen 
# Needs columns 'DATE_CAUGHT', 'STATION_ID', 'TAXONOMY_CODE_ID', 'SPECIMEN_NO'  

dat_specimens_to_add <- data.frame(
  DATE_CAUGHT = ymd_hms("2018-08-01 00:00:00"),    # from the already entered specimen data in Nivabase
  STATION_ID = 71087,
  TAXONOMY_CODE_ID = c(rep(8941, 3), rep(14089, 3)),   # 3 Zooplankton epilimnion, then 3 Mysis relicta  
  SPECIMEN_NO = c(1,2,3,1,2,3)
)

dat_specimens_to_add

```

### Make SQL   

* COPY-PASTE to SQL Developer (the code puts the SQL on the clipboard)  

```{r}

sql_list <- 1:nrow(dat_specimens_to_add) %>% 
  map_chr(make_sql_single_specimen, data = dat_specimens_to_add)
sql <- paste(sql_list, collapse = ";\n")
sql <- paste0(sql, ";\n")  
writeLines(sql, "clipboard")  # copies SQLs to clipboard - go to SQL Developer and paste

cat("Number of SQLs: \n")
length(sql_list)  # 9

cat("\nSample of SQLs: \n")
sql_list[1:3]

```

### Get the records we just added   

- Also checks that the number of new records is the same as no. of SQLs  
- Remember to **commit** in SQL developer first  
- Use this to add the last SPECIMEN_ID

```{r}

if (FALSE){

  # day_for_adding_records <- substr(lubridate::now(tzone = "UTC"), 1, 10)
  day_for_adding_records <- "2023-06-28"
  
  dat_specimens_added <- niRvana::get_nivabase_data(paste(
    "select DATE_CAUGHT, STATION_ID, TAXONOMY_CODE_ID, SPECIMEN_NO, SPECIMEN_ID",
    "from NIVADATABASE.BIOTA_SINGLE_SPECIMENS",
    "where TRUNC(ENTERED_DATE) = TO_DATE(", sQuote(day_for_adding_records), ", 'yyyy-mm-dd')", 
    " and ENTERED_BY = 'DHJ'")
  )
  dat_specimens_added
  
  cat("Number of retrieved records: ")
  n_records2 <- nrow(sed_methods_added_to_database)
  n_records2
  
}

# Use outpput from the above to add SPECIMEN_ID    
lookup_id <- data.frame(
  ID = c(
    paste0("ZM-", 1:3),        # zooplankton
    paste0("MM-", 1:3),        # Mysis
    paste0("KM-M-", 1:10),     # Krøkle = Osmerus eperlanus
    paste0("LM-M-", 1:10),     # Lågåsild = Coregonus albula
    paste0("ØM-M-", 1:15),     # Ørret = Salmo trutta, Mjøsa
    paste0("ØF-M-", 1:10)      # Ørret = Salmo trutta, Femunden
    ),
  SAMPLE_ID = c(248738:248739, 248787,
                248740:248741, 248788,
                248742:248751,
                248752:248761,
                248762:248776,
                248777:248786),
  SPECIMEN_ID = c(334032:334034,    # zooplankton
                  334035:334037,    # Mysis
                  333955:333964,
                  333965:333974,
                  333975:333989,
                  333990:333999)
) %>%
  mutate(ID = fct_inorder(factor(ID)))

```


## 5. Add to BIOTA_SAMPLES_SPECIMENS  

### Make SQL   

* COPY-PASTE to SQL Developer (the code puts the SQL on the clipboard)  

```{r}

# make_sql_samples_specimens

sql_list <- 1:nrow(lookup_id) %>% 
  map_chr(make_sql_samples_specimens, data = lookup_id)
sql <- paste(sql_list, collapse = ";\n")
sql <- paste0(sql, ";\n")  
writeLines(sql, "clipboard")  # copies SQLs to clipboard - go to SQL Developer and paste

cat("Number of SQLs: \n")
length(sql_list)  # 9

cat("\nSample of SQLs: \n")
sql_list[1:3]

```
