---
title: "400_VMexport_Urbanfjord_2024"
format: html
editor: visual
---

## For filenames  
```{r}

program <- "urbanfjord"
sampling_year <- 2024

```

## 1. Packages and functions

```{r}

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)
library(jsonlite)
library(glue)      # for glue
library(datapasta)
library(stringdist)    # for functions in '400_parameter_matching_functions.R'
# library(fuzzyjoin)   # regex_full_join

source("400_parameter_matching_functions.R")   # for parameter matching, see part 4

# source("../Milkys/991_Vannmiljo_snippets_functions.R")
# source("../Milkys/992_Vannmiljo_Urban_fjord_functions.R")
# source("../Milkys/994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```

## 2. Read data

```{r}

df_water_1 <- readRDS('Data/200_dat_water_Urban_2024_2025-06-17.rds')
df_biota_1 <- readRDS('Data/200_dat_biota_Urban_2024_2025-06-17.rds')
df_stations_1 <- readRDS('Data/200_dat_stations_Urban_2024_2025-06-17.rds')

```

## 3. Add Vannmiljø codes for stations

### Add VANNKATEGORI 1

-   Take VANNKATEGORI from the AqM log for existing stations

```{r}

df_stations_1 <- df_stations_1 %>% 
  mutate(
    Kilde_id = paste0("NIVA@", STATION_ID))

ids <- df_stations_1$Kilde_id

# Read the VANNMILJO log in AqM
#   (how did we find this? see Appendix)  
df_log_all <- get_nivabase_selection(
  "*",
  "LOGG_LOKALITETER", 
  "VANNLOKALITETID",
  ids, 
  owner = "VANNMILJO", values_are_text = TRUE)

# summarize by taking the last log for each station  
df_log_stations <- df_log_all %>% 
  arrange(desc(ENTERED_DATE)) %>% 
  group_by(VANNLOKALITETID) %>% 
  summarize(across(everything(), .fns = first))

# Add existing VANNKATEGORI to the data  
df_stations_2 <- df_stations_1 %>% 
  left_join(
    df_log_stations %>% select(VANNLOKALITETID, VANNKATEGORI), 
    by = join_by(Kilde_id == VANNLOKALITETID))

# Check
table(addNA(df_stations_2$VANNKATEGORI))
# R = river  
# U = uncategorized (used for storm water)

sel_lacking <- is.na(df_stations_2$VANNKATEGORI)

if (sum(sel_lacking) > 0){
  warning(sum(sel_lacking), " stations lacking VANNKATEGORI (use next chunk)")
}

```

### Add VANNKATEGORI 2

-   Add VANNKATEGORI for stations without it (new stations)\
-   

```{r}

# for making the code blow, first run this
df <- df_stations_2[sel_lacking,] %>% 
  mutate(VANNKATEGORI_new = "U") %>%    # this is just temporary - will be corrected in the tribble code below
  select(Kilde_id, STATION_CODE, STATION_NAME, VANNKATEGORI_new) 

# use datapasta to start code of data shown below
# run this
#   datapasta::tribble_format(df)
# and then Ctrl + V

# HARD_CODED FOR 2024 specifically
# See Figure 1B in the report to set df_for_vannkategori
df_for_vannkategori <- tibble::tribble(
  ~Kilde_id,              ~STATION_CODE,                                     ~STATION_NAME, ~VANNKATEGORI_new,
      "NIVA@74314",            "ALNA Evert. 1",                         "Alna River Evertebrat 1",           "R",
      "NIVA@74315",              "Alna Mose 2",                               "Alna River Moss 2",           "R",
      "NIVA@74316",            "ALNA Evert. 2",                         "Alna River Evertebrat 2",           "R",
      "NIVA@74317",              "Alna Mose 3",                               "Alna River Moss 3",           "R",
      "NIVA@74318",            "ALNA Evert. 3",                         "Alna River Evertebrat 3",           "R",
      "NIVA@74637",        "Kværner (Kildesp)",                                         "Kværner",           "R",
      "NIVA@74638",     "Smalvollv. (Kildesp)",                                   "Smalvollveien",           "U",
      "NIVA@74639",    "Trosterudb. (Kildesp)",                                 "Trosterudbekken",           "U",
      "NIVA@74640",       "Breivoll (Kildesp)",                             "Breivoll (Alnaelva)",           "R",
      "NIVA@74641",       "Strømsv. (Kildesp)",                                 "Strømsveien 222",           "U",
      "NIVA@74642", "Verks.Furul.v. (Kildesp)",                      "Verkseier Furulunds vei 9B",           "U",
      "NIVA@74643",        "Vollav. (Kildesp)", "Vollaveien 20B. Enden av overvannsrør ut i Alna",           "U",
      "NIVA@74644",     "Kjelsrudf. (Kildesp)",  "Avrenning Kjelsrudfyllingene. Rett før Kulvert",           "U",
      "NIVA@74645",         "Brubak (Kildesp)",                               "Brubak (Alnaelva)",           "R",
      "NIVA@74646",       "Stansev. (Kildesp)",                                  "Stanseveien 40",           "U",
      "NIVA@74647",     "Julsbergb. (Kildesp)",             "Julsbergbekken. Utløp etter kulvert",           "U",
      "NIVA@74648",    "Tokerudb. 1 (Kildesp)",              "Tokerudbekken, utløp etter kulvert",           "U",
      "NIVA@74649",    "Tokerudb. 2 (Kildesp)",           "Tokerudbekken, oppstrøms bekkelukking",           "U"
  )

df_stations_3 <- df_stations_2 %>% 
  left_join(
    df_for_vannkategori %>% select(Kilde_id, VANNKATEGORI_new), 
    by = join_by(Kilde_id)) %>% 
  mutate(
    VANNKATEGORI = case_when(
      is.na(VANNKATEGORI_new) ~ VANNKATEGORI,
      !is.na(VANNKATEGORI_new) ~ VANNKATEGORI_new))

table(addNA(df_stations_3$VANNKATEGORI))

# check
# df_stations_3 %>% select(Kilde_id, STATION_CODE, STATION_NAME, VANNKATEGORI, VANNKATEGORI_new) %>%  View("kategori") 

```

### UTM zone

```{r}

table(df_stations_3$EPSG_SYS)

lookup_utmzone <- tibble::tribble(
  ~EPSG_SYS,   ~UtmSone,
  32630,       30,
  32631,       31,
  32632,       32,
  32633,       33,
  32634,       34,
  32635,       35,
  32636,       36
)

df_stations_4 <- df_stations_3 %>% 
  left_join(lookup_utmzone, by = join_by(EPSG_SYS), relationship = "many-to-one")


```

### For Vannmiljø

```{r}

df_stations_vm <- df_stations_4 %>% 
  rename(
    Navn = STATION_NAME,
    Latitude = LATITUDE,
    Longitude = LONGITUDE,
    Vannkategori = VANNKATEGORI) %>% 
  mutate(
    Beskrivelse = "",
    DatoOpprettet	= "", 
    Hoyde = "",
    UtmNord	= ifelse(is.na(UtmSone), NA, Y),
    UtmOst	= ifelse(is.na(UtmSone), NA, X)) %>% 
  select(
    Kilde_id, Navn, Beskrivelse, DatoOpprettet, Hoyde, Latitude, Longitude, UtmSone, UtmNord, UtmOst, Vannkategori)


```

## 4. Add Vannmiljø codes for water

### ID codes

```{r}

df_water_2 <- df_water_1 %>%
  mutate(
    Vannlok_kode = paste0("NIVA@", STATION_ID),       # note: called 'Kilde_id' in station file (LOK)  
    ID_lokal = paste0("NIVA@WC@", VALUE_ID)) %>%      # note: WATER SAMPLE version
  # add VANNKATEGORI - this is needed for water samples on order to set Provetakmetode + Medium  
  left_join(
    df_stations_4 %>% select(Kilde_id, VANNKATEGORI), 
    by = join_by(Vannlok_kode == Kilde_id ), 
    relationship = "many-to-one") 
  

```

### Prøvetakmetode + medium 1

-   Set sample type\
-   Note that most substances in most locations are divided in water and particle phase\
-   Exceptions:
    -   Klorheksidin (Chlorhexidine) is in addition taken in unfiltered water (for source tracing)
    -   Diklormetan and Triklormetan (kloroform) (Dichloromethane and trichloromethane) also in unfiltered water for source tracing

```{r}

df_water_3 <- df_water_2 %>%
  mutate(
    NAME_nivabase = NAME,
    NAME = sub("-P", "", NAME_nivabase, fixed = TRUE),
    Particle_phase = grepl("-P", NAME_nivabase)
  ) %>% 
  # Also need to mark both particle and water fraction with 'Sample_filtered' 
  # since this needs an extra 
  group_by(
    WATER_SAMPLE_ID, NAME) %>% 
  mutate(
    Particle_phase_n = sum(Particle_phase)) %>% 
  ungroup() %>% 
  mutate(
    Sample_filtered = Particle_phase_n > 0) 

# View(df_water_3)
# df_water_3 %>% filter(NAME == "PFOSA") %>% View()

tabulate_samples <- function(param){
  cat("----", param, "----\n")
  df_water_3 %>% 
    filter(NAME == param) %>%
    xtabs(~Sample_filtered + Particle_phase, .) %>% 
    print()
  cat("\n")
}

# All samples filtered, both phases
tabulate_samples("PFOS")
# Some samples filtered, some not
tabulate_samples("Klorheksidin")
# No samples filtered
tabulate_samples("Triklormetan (kloroform)")
```

### Prøvetakmetode + medium 2

-   Set sample type codes

```{r}

lookup_vm_medium <- fromJSON(
  "https://vannmiljoapi.miljodirektoratet.no/api/Public/GetMediumList?filter=")
# lookup_vm_medium

lookup_vm_samplingmethod <- fromJSON(
  "https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSamplingMethodList?filter=")

# NOTE: this is very Urban Fjord-specific  
# In particular, STATION_ID 69644 is Bekkelaget Renseanlegg (Bekkelaget WWTP)  
df_water_4 <- df_water_3 %>% 
  mutate(
    Medium_id = case_when(
      Particle_phase ~ "PA"       ,      # Partikler
      # STATION_ID %in% 69644 ~ "VAR",   # Avløpsvann renset   # NOTE: very Urban Fjord-specific" 
      VANNKATEGORI %in% "R" ~ "VF",      # Ferskvann (river water)
      VANNKATEGORI %in% "U" ~ "VOV"),    # Overvann
    Provetakmetode_id = case_when(
      Particle_phase ~ "NS-ISO 5667-17:2008D",                           # Partikler: Filtrering volumbasert
      !Particle_phase & Sample_filtered ~ "UKJENT",                      # Used for the water phase of filtered samples 
      !Particle_phase & !Sample_filtered ~ "NS-EN ISO 5667-6:2016")      # Ferskvann (elver og bekker)
    ) %>% 
  left_join(
    lookup_vm_medium %>% select(MediumID, Medium_name = Name), 
    by = join_by(Medium_id == MediumID)) %>% 
  left_join(
    lookup_vm_samplingmethod %>% select(SamplingMethodID, Provetakmetode_name = Name), 
    by = join_by(Provetakmetode_id == SamplingMethodID))

xtabs(~addNA(Medium_id) + addNA(Provetakmetode_id), df_water_4)
xtabs(~addNA(Medium_name) + addNA(Provetakmetode_name), df_water_4)

```

### Units 1: Fix units

-   Fix units for metals in particle phase\
-   See script 200, issue no. 3

```{r}

# UNIT = NONE 
table(addNA(df_water_4$UNIT))

tabulate_samples("Yttrium")

# example
df_water_4 %>% 
  filter(NAME %in% "Yttrium" & Sample_filtered) %>%
  select(WATER_SAMPLE_ID, Particle_phase, Sample_filtered, NAME, VALUE, UNIT)

df_water_affected_samples <- df_water_4 %>% 
  filter(Sample_filtered) %>%
  select(WATER_SAMPLE_ID, Particle_phase, NAME, UNIT) %>% 
  pivot_wider(names_from = Particle_phase, values_from = UNIT, names_prefix = "Particle_") %>% 
  filter(Particle_TRUE %in% "NONE")

xtabs(~WATER_SAMPLE_ID + NAME, df_water_affected_samples)

df_water_4b <- df_water_4 %>% 
  left_join(
    df_water_affected_samples %>%
      select(WATER_SAMPLE_ID, NAME, UNIT_new = Particle_FALSE), 
    by = join_by(WATER_SAMPLE_ID, NAME), 
    relationship = "many-to-one")

# check  
df_water_4b %>% 
  filter(NAME %in% "Yttrium" & Sample_filtered) %>%
  select(WATER_SAMPLE_ID, Particle_phase, Sample_filtered, NAME, VALUE, UNIT, UNIT_new)

df_water_5 <- df_water_4b %>% 
  mutate(
    UNIT = case_when(
      UNIT %in% "NONE" ~ UNIT_new,
      TRUE ~ UNIT)
  ) %>% 
  select(
    -UNIT_new)
  
```

### Units 2: Add VM code

* Note: all metals codes will be changed to ug/L later - see 'Units 3'  

```{r}

lookup_vm_unit_all <- fromJSON(
  "https://vannmiljoapi.miljodirektoratet.no/api/Public/GetUnitList?filter=")
# lookup_vm_unit

table(df_water_5$UNIT)

# for starting on code below
lookup_vm_unit_all %>% 
  filter(UnitID %in% c(8,43,58)) %>% 
  select(UnitID, Name, Description) %>%
  mutate(UNIT = "") %>% 
  datapasta::tribble_format()

lookup_vm_unit <- tibble::tribble(
  ~UnitID,  ~Name,          ~Description, ~UNIT,
       8L, "µg/l", "mikrogram per liter",    "NG_P_ML",
       8L, "µg/l", "mikrogram per liter",    "NG_P_ML_2",
      43L, "mg/l", "milligram per liter",    "MG_P_L",
      58L, "ng/l",  "nanogram per liter",    "NG_P_L_2",
      58L, "ng/l",  "nanogram per liter",    "NG_P_L")

df_water_6 <- df_water_5 %>% 
  left_join(
    lookup_vm_unit %>% select(UNIT, UnitID, Unit_VM = Name),
    by = join_by(UNIT))

xtabs(~addNA(Unit_VM) + addNA(UNIT), df_water_6)

```


### Parameters 1 - find matches   

* From NAME (Nivabase) to Vannmiljø parameter name  
* Strategy:
    - match NAME (Dataset1 in part a) to "Abbreviation" (Dataset2 in part a) in the report Appendix  
    - from the report Appendix, get CAS number (manually edited for a fes cases)
    - from the CAS number, get Vannmiljø parameter name

#### a. Perform matching 

* using functions made by Claude, stored in '400_parameter_matching_functions.R'

```{r}

lookup_substances_cas_all <- readxl::read_excel(
  "Input_data/Lookup_files/Urbanfjord_2024_parameter_list.xlsx") %>% 
  filter(!is.na(Abbreviation))

# 'norwegian_english_dict ' should be checked manually
# View(norwegian_english_dict)

dataset1 <- df_water_6 %>% 
  distinct(NAME, METHOD_ID)

dataset2 <- lookup_substances_cas_all

# Perform the matching
matching_results <- match_datasets(dataset1, dataset2, norwegian_english_dict)

# Display summary statistics
cat("=== MATCHING SUMMARY ===\n")
cat("Total items in Dataset 1:", nrow(dataset1), "\n")
cat("Total items in Dataset 2:", nrow(dataset2), "\n")
cat("Total matches found:", sum(matching_results$Match_Type != "No Match Found"), "\n")
cat("Unmatched items:", sum(matching_results$Match_Type == "No Match Found"), "\n\n")

```

#### b. Breakdown of match statistics

* Breakdown of match type and quality 

```{r}

# Display match type breakdown
match_type_summary <- table(matching_results$Match_Type)
cat("=== MATCH TYPE BREAKDOWN ===\n")
print(match_type_summary)
cat("\n")

# Display match quality breakdown
match_quality_summary <- table(matching_results$Match_Quality)
cat("=== MATCH QUALITY BREAKDOWN ===\n")
print(match_quality_summary)
cat("\n")

```
#### c. Unmatched parameters

```{r}

unmatched_items <- matching_results[matching_results$Match_Type == "No Match Found", ]
# str(unmatched_items)

if (F){
  # this is used to make the code below
  # run the code and use Ctrl+V to paste the resukt below
  # then change "-" manually in the code
  unmatched_items %>% 
    select(NAME = Dataset1_NAME, METHOD_ID = Dataset1_METHOD_ID) %>% 
    mutate(Abbreviation = "-") %>% 
    datapasta::tribble_format()
}

df_manual <- tibble::tribble(
                                       ~ NAME,          ~METHOD_ID,          ~Abbreviation,
#                                "Diklormetan",               41601,                  "DCM",
                                 "Bisfenol P",               40274,                  "Bisphenol P",
                                       "BnBP",               40310,                  "BnBP (DBOEP)",
                                       "BMEP",               40325,                  "BMEP (DMEP)",
                 "L3 (octamethyltrisiloxane)",               40413,                  "L3",
         "D6 (dodekametylsykloheksasiloksan)",               40414,                  "D6",
           "D5 (dekametylsyklopentasiloksan)",               40416,                  "D5",
  "M3T(Ph) (tris(trimetylsiloksy)fenylsilan)",               40418,                  "M3T (Ph)",
             "L5 (dodecamethylpentasiloxane)",               40419,                  "L5",
           "D4 (oktametylsyklotetrasiloksan)",               40420,                  "D4",
               "L4 (decamethyltetrasiloxane)",               40421,                  "L4",
                 "L3 (octamethyltrisiloxane)",               40936,                  "L3",
                                       "BnBP",               40942,                  "BnBP (DBOEP)",
                                 "Bisfenol P",               41036,                  "Bisphenol P",
         "D6 (dodekametylsykloheksasiloksan)",               41050,                  "D6",
           "D5 (dekametylsyklopentasiloksan)",               41072,                  "D5",
  "M3T(Ph) (tris(trimetylsiloksy)fenylsilan)",               41094,                  "M3T (Ph)",
             "L5 (dodecamethylpentasiloxane)",               41101,                  "L5",
                                       "BMEP",               41111,                  "BMEP (DMEP)",
           "D4 (oktametylsyklotetrasiloksan)",               41114,                  "D4",
               "L4 (decamethyltetrasiloxane)",               41127,                  "L4",
           "L6 (Tetradecamethylhexasiloxane)",               41603,                  "L6",
           "L7 (Hexadecamethylheptasiloxane)",               41608,                  "L7",
            "L8 (Octadecamethyloctasiloxane)",               41609,                  "L8",
           "L6 (Tetradecamethylhexasiloxane)",               41610,                  "L6",
           "L7 (Hexadecamethylheptasiloxane)",               41630,                  "L7",
            "L8 (Octadecamethyloctasiloxane)",               41636,                  "L8",
           "L7 (Hexadecamethylheptasiloxane)",               41683,                  "L7",
            "L8 (Octadecamethyloctasiloxane)",               41689,                  "L8"
  )

```


### Parameters 2: make lookup file with CAS    

```{r}

lookup_substances_1a <- matching_results[matching_results$Match_Type != "No Match Found", ] %>%
  rename(
    NAME = Dataset1_NAME, METHOD_ID = Dataset1_METHOD_ID, 
    Substances = Dataset2_Substances, Abbreviation = Dataset2_Abbreviation, Group = Dataset2_Group
  )

lookup_substances_1b <- df_manual %>%
  left_join(lookup_substances_cas_all %>% select(Abbreviation, Substances, Group), 
            by = join_by(Abbreviation), 
            relationship = "many-to-one")

# Check
# head(lookup_substances_1a, 3)
# head(lookup_substances_1b, 3)

# check manually:
# View(matching_results_2b)

lookup_substances_1 <- bind_rows(lookup_substances_1a, lookup_substances_1b)

lookup_substances_2 <- lookup_substances_1 %>% 
  left_join(lookup_substances_cas_all, 
            by = join_by(Abbreviation, Substances, Group), 
            relationship = "many-to-one")

```


### Parameters 2: make lookup file with Vannmiljo names  

```{r}

lookup_vm_substances_all <- fromJSON(
  "https://vannmiljoapi.miljodirektoratet.no/api/Public/GetParameterList?filter=")

lookup_vm_substances <- lookup_vm_substances_all %>% 
  add_count(CASnr) %>% 
  filter(n == 1) %>% 
  select(CASnr, ParameterID, VM_Name = Name, VM_Description = Description)

lookup_substances_3 <- lookup_substances_2 %>% 
  left_join(
    lookup_vm_substances,
    by = join_by(CAS_VM == CASnr), 
    relationship = "many-to-one")

```

### Parameters 3: add ParameterID manually from VM 

```{r}

if (F){

  # check
  lookup_substances_3 %>% 
    filter(is.na(ParameterID)) %>% 
    View("lacking ParameterID")
  
  # this is used to make the code below
  # run the code and use Ctrl+V to paste the resukt below
  # then change "-" manually in the code
  lookup_substances_3 %>% 
    filter(is.na(ParameterID)) %>% 
    select(NAME, Substances, Abbreviation, Group, ParameterID) %>% 
    mutate(ParameterID = "-") %>% 
    datapasta::tribble_format()
}

lookup_substances_3_manual_1 <- tibble::tribble(
           ~NAME,                       ~Substances,  ~Abbreviation,                          ~Group, ~ParameterID,
  "Klorheksidin",                   "Chlorhexidine", "Klorheksidin",         "Pesticides/Fungicides",          "CLHEX",
    "Galaxolide",                      "Galaxolide",         "HHCB",                         "Musks",          "HHCB",
    "DADMAC-C12",       "Didodecyldimethylammonium",   "DADMAC-C12", "Quaternary ammonium compounds",          "DADMAC-C12",
    "DADMAC-C14",    "Dimethylditetradecylammonium",   "DADMAC-C14", "Quaternary ammonium compounds",          "DADMAC-C14",
    "DADMAC-C16",     "Dihexadecyldimethylammonium",   "DADMAC-C16", "Quaternary ammonium compounds",          "DADMAC-C16",
    "DADMAC-C18",     "Dimethyldioctadecylammonium",   "DADMAC-C18", "Quaternary ammonium compounds",          "DADMAC-C18",
      "ATAC-C16",      "Hexadecyltrimethylammonium",     "ATAC-C16", "Quaternary ammonium compounds",          "ATAC-C16",
      "ATAC-C18",      "Trimethyloctadecylammonium",     "ATAC-C18", "Quaternary ammonium compounds",          "ATAC-C18",
       "BAC-C10",     "Benzyldimethyldecylammonium",      "BAC-C10", "Quaternary ammonium compounds",          "BAC-C10",
       "BAC-C16", "Benzyldimethylhexadecylammonium",      "BAC-C16", "Quaternary ammonium compounds",          "BAC-C16",
  )

# check
# lookup_substances_3_manual_1 %>% 
#   filter(NAME != ParameterID)

lookup_substances_3_manual <- lookup_substances_3_manual_1 %>% 
  left_join(
    lookup_vm_substances %>% select(CASnr, ParameterID, VM_Name, VM_Description),
    by = join_by(ParameterID))

```

### Parameters 4: Final lookup table  

```{r}

lookup_substances <- bind_rows(
  lookup_substances_3 %>% filter(!is.na(ParameterID)),
  lookup_substances_3_manual
)

check <- is.na(lookup_substances$ParameterID)

if (sum(check) > 0){
  stop("Still lacking some ParameterID values")
}

check <- lookup_substances %>% 
  add_count(METHOD_ID, NAME) %>% 
  filter(n > 1) %>%
  arrange(METHOD_ID, NAME) %>% 
  select(NAME, METHOD_ID, Group, ParameterID, CAS_VM)

if (nrow(check) > 0){
  stop("Some NAME, METHOD_ID occur several times")
}

if (FALSE){
  readr::write_csv2(lookup_substances, "Input_data/Lookup_files/lookup_substances_urbanfjord_water_2024.csv")
  writexl::write_xlsx(lookup_substances, "Input_data/Lookup_files/lookup_substances_urbanfjord_water_2024.xlsx")
}

```

### Parameters 5: add ParameterID to data  
```{r}

lookup_substances2 <- lookup_substances %>% 
  distinct(NAME, ParameterID)

df_water_7 <- df_water_6 %>% 
  left_join(lookup_substances2, 
            by = join_by(NAME),
            relationship = "many-to-one")

check <- df_water_7 %>% 
  filter(is.na(ParameterID)) %>% 
  distinct(NAME) %>% 
  select(NAME)

if (nrow(check) > 0){
  stop("some lacking!")
}

```

### Units 3: Change metal units to mg/L  

* all metals should be µg/l (for water) or mg/kg (for biota)  
    - see mail from Dag Rosland 27.06.2025: "En del av metall-dataene er oppgitt med feil enhetsID = 7 ( µg/kg v.v.) for biotamediene ('BH', 'BL', 'BM' og 'BPV') og feil enhetsID = 58 (ng/l) for vann- og partikkel-mediene ('VF', 'VOV' og 'PA'). Korrekt enhetsID skal være hhv. 42 (mg/kg v.v.) og 8  (µg/l). Fint om du kan korrigere enhetsID i vedlagte Excel-ark og sjekke om verdiene er oppgitt i de nevnte enhetene.


```{r}

if (FALSE){
  # check with Dag Rosland mail 
  # These have wrong units according to Dag R.:
  metals_to_change <- c("DY","ER","EU","GD","HG", "HO","LA","LU","ND","PR","SB","SC","SM","TB","TM","YB")  
  # Check that all are given as Group = metals:
  lookup_substances %>% 
    filter(ParameterID %in% metals_to_change) %>% 
    xtabs(~addNA(Group), .)
}

metals <- subset(lookup_substances, Group %in% "Metals")$ParameterID |> unique()
sel_rows <- with(df_water_7, ParameterID %in% metals & UnitID %in% 58)   # UnitID = 58: ng/L
sum(sel_rows)
# table(df_water_7$ParameterID[sel_rows])

df_water_8 <- df_water_7
df_water_8$VALUE <- df_water_7$VALUE/1000  
df_water_8$UnitID <- 8                      # UnitID = 8: µg/L  

```

### Data for Vannmiljo 1

```{r}

df_water_vm_1 <- df_water_8 %>% 
  rename(
    Parameter_id = ParameterID,
    Verdi = VALUE,
    Enhet_id = UnitID
  ) %>% 
  mutate(
    Aktivitet_id = "MIUR",
    Oppdragsgiver = "Miljødirektoratet",
    Oppdragstaker = "NIVA",
    LatinskNavn_id = as.numeric(NA),
    Filtrert_prove = "Nei",
    Analysemetode_id = case_when(
      Parameter_id %in% "SCB7" ~ "NS-EN ISO 6468",
      TRUE ~ "UKJENT"),
    Operator = case_when(
      is.na(FLAG1) ~ "=",
      TRUE ~ "<"),
    Provenr = as.character(NA),
    Tid_provetak = floor_date(SAMPLE_DATE, unit = "day") %>% as.character(),
    Ovre_dyp = 0,
    Nedre_dyp = 0,
    DybdeEnhet = "m",
    Ant_verdier = NA,
    Kommentar = "",
    Deteksjonsgrense = NA,
    Kvantifiseringsgrense = NA
  )

table(addNA(df_water_8$FLAG1))
table(addNA(df_water_vm_1$Operator))
table(addNA(df_biota_vm_2$Operator))

```

### Data for Vannmiljo 2

```{r}

fn <- "Data_to_Vannmiljo/urban_vm/REG_water_2019-2021_kovirbv.xlsx"
# readxl::excel_sheets(fn)
df_example <- readxl::read_excel(fn, sheet = "REG_water_2019-2021_kovirbv")

lookup_kommentar <- df_example %>% 
  distinct(Parameter_id, Kommentar)

check <- lookup_kommentar %>% 
  add_count(Parameter_id) %>% 
  filter(n > 1) %>%
  arrange(Parameter_id)

lookup_kommentar <- lookup_kommentar %>% 
  filter(!is.na(Kommentar))

df_water_vm_2 <- df_water_vm_1 %>% 
  select(-Kommentar) %>% 
  left_join(lookup_kommentar, by = "Parameter_id")

```

### Data for Vannmiljo 3
```{r}

# names(df_example) %>% dput()

vars <- c("ID_lokal", "Vannlok_kode", "Aktivitet_id", "Oppdragsgiver", 
          "Oppdragstaker", "Parameter_id", "Medium_id", "LatinskNavn_id", 
          "Provetakmetode_id", "Filtrert_prove", "Analysemetode_id", "Operator", 
          "Verdi", "Enhet_id", "Provenr", "Tid_provetak", "Ovre_dyp", "Nedre_dyp", 
          "DybdeEnhet", "Ant_verdier", "Kommentar", "Deteksjonsgrense", 
          "Kvantifiseringsgrense")

df_vm_water <- df_water_vm_2[vars]

```


## 5. Add Vannmiljø codes for biota

### Species

-   All
    -   https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSpeciesList?filter
-   Example
    -   https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSpeciesList?filter=Insecta
-   LATIN_NAME = 'NA' hardcoded to 89 (Insecta)

```{r}


# test
# fromJSON("https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSpeciesList?filter=Insecta")

lookup_vm_species <- fromJSON("https://vannmiljoapi.miljodirektoratet.no/api/Public/GetSpeciesList?filter=")

df_biota_2 <- df_biota_1 %>% 
  left_join(
    lookup_vm_species %>% select(VitenskapligNavn, LatinskNavnID), 
    by = join_by(LATIN_NAME == VitenskapligNavn), 
    relationship = "many-to-one")

table(addNA(df_biota_2$LatinskNavnID))

sel <- is.na(df_biota_2$LatinskNavnID)
sum(sel)
# table(addNA(df_biota_2$LATIN_NAME[sel]))

if (sum(sel) > 0){
  warning("Species hard-coded where not found")
}

# HARD-CODED - see 8a above 
df_biota_2$LatinskNavnID[sel] <- 89            # HARD-CODED FOR 2024 specifically
df_biota_2$LATIN_NAME[sel] <- "Insecta"        # HARD-CODED FOR 2024 specifically

# Check
table(addNA(df_biota_2$LATIN_NAME))

```

### Tissue

```{r}

table(df_biota_2$TISSUE_NAME)

sel <- df_biota_2$TISSUE_NAME %in% "Annet"

if (sum(sel) > 0){
  warning("Tissue hard-coded where not found")
  df_biota_2$TISSUE_ID[sel] <- 104                   # HARD-CODED FOR 2024 specifically
  df_biota_2$TISSUE_NAME[sel] <- "Plantevev"         # HARD-CODED FOR 2024 specifically  
  
}

df_convmedium <- get_nivabase_data(
  "select TISSUE_ID, MEDIUM_ID, DATATYPE_ID, STATION_TYPE_ID from VANNMILJO.CONV_MEDIUM")

df_biota_3 <- df_biota_2 %>% 
  left_join(
    df_convmedium %>% select(TISSUE_ID, MEDIUM_ID), 
    by = join_by(TISSUE_ID), 
    relationship = "many-to-one")

sel <- is.na(df_biota_3$MEDIUM_ID)
sum(sel)

if (sum(sel) > 0){
  stop("VN tissues lacking")
} else 
{
  message("All VM tissues OK")
}


```


### Units 1: Make lookup

```{r}

table(addNA(df_biota_3$UNIT))

# df_biota_3 %>% 
#   filter(UNIT == "NONE") %>% View("unit NONE")

lookup_vm_unit_all <- fromJSON(
  "https://vannmiljoapi.miljodirektoratet.no/api/Public/GetUnitList?filter=")

# check and search for "våt" in Description field
# View(lookup_vm_unit_all)

if (FALSE){
  # for starting on code below
  lookup_vm_unit_all %>% 
    filter(UnitID %in% c(1,2,7,42)) %>% 
    select(UnitID, Name, Description) %>%
    mutate(UNIT = "-") %>% 
    datapasta::tribble_format()
}

lookup_vm_unit <- tibble::tribble(
  ~UnitID,        ~Name,                         ~Description, ~UNIT,
       1L,          "%",                            "prosent",   "PERCENT",
       2L,     "% v.v.",            "prosent på våtvektbasis",   "PERCENTW_W",
       7L, "µg/kg v.v.", "mikrogram per kilo på våtvektbasis",   "NG_P_G",
      42L, "mg/kg v.v.", "milligram per kilo på våtvektbasis",   "MG_P_KG"
  )



```
### Units 2: Add VM code

```{r}

df_biota_4 <- df_biota_3 %>% 
  left_join(lookup_vm_unit %>% select(UNIT, UnitID), 
            by = join_by(UNIT),
            relationship = "many-to-one")

```


### Parameters: add ParameterID to data  

```{r}

# lookup_substances <- readxl::read_excel("Input_data/Lookup_files/lookup_substances_urbanfjord_water_2024.xlsx")

lookup_substances2 <- lookup_substances %>% 
  distinct(NAME, ParameterID)

check <- lookup_substances2 %>% 
  add_count(NAME) %>% 
  filter(n > 1) %>%
  arrange(NAME)

if (nrow(check) > 0){
  stop("NAME not unique")
}

df_biota_5 <- df_biota_4 %>% 
  left_join(lookup_substances2 %>% select(NAME, ParameterID), 
            by = join_by(NAME),
            relationship = "many-to-one") %>% 
  mutate(
    ParameterID = case_when(
      NAME %in% "Delta 13C" ~ "D13C",
      NAME %in% "Delta 15N" ~ "D15N",
      NAME %in% "Fettinnhold" ~ "LIPID",
      NAME %in% "L9 (Eicosamethylnonasiloxane)" ~ "L9",
      NAME %in% "L10 (Docosamethyldecasiloxane)" ~ "L10",
      TRUE ~ ParameterID)
  ) %>% 
  filter(
    !NAME %in% c("W% C/N", "% C", "% N")
  )

check <- df_biota_5 %>% 
  filter(is.na(ParameterID)) %>% 
  distinct(NAME) %>% 
  select(NAME)

if (nrow(check) > 0){
  stop("some lacking!")
}


```

### Add Medium ID

```{r}

lookup_vm_medium <- fromJSON(
  "https://vannmiljoapi.miljodirektoratet.no/api/Public/GetMediumList?filter=")
# lookup_vm_medium

lookup_vm_species %>% 
  filter(LatinskNavnID %in% c(27, 126, 16496, 108169))

lookup_vm_medium

if (FALSE){
  table(addNA(df_biota_5$TISSUE_NAME))
  table(addNA(df_biota_5$LATIN_NAME))
  
  lookup_vm_medium %>% 
    filter(Name %in% c("Biota bløtdeler", "Biota helkropp", "Biota lever", "Biota muskelvev", "Biota plantevev")) %>% 
    select(MediumID, Name) %>%
    mutate(TISSUE_NAME = "-",
           LATIN_NAME = "-") %>% 
    datapasta::tribble_format()
}

lookup_medium_biota <- tibble::tribble(
  ~MediumID,             ~Name, ~TISSUE_NAME, ~LATIN_NAME,
       "BB", "Biota bløtdeler",          "WHole soft body",       "Mytilus edulis",
       "BB", "Biota bløtdeler",          "WHole soft body",       "Pandalus borealis",
       "BH",  "Biota helkropp",          "Hel organisme",         "Insecta",
       "BH",  "Biota helkropp",          "Hel organisme",         "Polychaeta (Klasse)",
       "BH",  "Biota helkropp",          "Hel organisme",         "Euphausiacea (Orden)",
       "BL",     "Biota lever",          "Lever",         "Salmo trutta",
       "BM", "Biota muskelvev",          "Muskel",         "Salmo trutta",
      "BPV", "Biota plantevev",          "Plantevev",         "Fontinalis antipyretica"
)

df_biota_6 <- df_biota_5 %>% 
  left_join(lookup_medium_biota %>% select(MediumID, TISSUE_NAME, LATIN_NAME),
            by = join_by(TISSUE_NAME, LATIN_NAME), relationship = "many-to-one"
  )

```
### Units 3: Change metal units to mg/kg  

* all metals should be mg/L (for water) or mg/kg (for biota)  
* similar to "Units 3" for water, see above, except that here we operate on per kg basis   
* enhetsID = 7 ( µg/kg v.v.) to be changed to 42 (mg/kg v.v.)

```{r}

metals <- subset(lookup_substances, Group %in% "Metals")$ParameterID |> unique()
sel_rows <- with(df_biota_6, ParameterID %in% metals & UnitID %in% 7)   # UnitID = 7: µg/kg
sum(sel_rows)
# table(df_water_7$ParameterID[sel_rows])

df_biota_7 <- df_biota_6
df_biota_7$VALUE <- df_biota_6$VALUE/1000  
df_biota_7$UnitID <- 42                      # UnitID = 42: mg/kg v.v.  

```

### Data for Vannmiljo 1

```{r}

df_biota_vm_1 <- df_biota_7 %>% 
  rename(
    Parameter_id = ParameterID,
    Verdi = VALUE,
    Enhet_id = UnitID,
    Medium_id = MediumID,                # different from water
    LatinskNavn_id = LatinskNavnID       # different from water
  ) %>% 
  mutate(
    Aktivitet_id = "MIUR",
    Oppdragsgiver = "Miljødirektoratet",
    Oppdragstaker = "NIVA",
    Provetakmetode_id = "UKJENT",    # different from water
    Filtrert_prove = "Nei",
    Analysemetode_id = case_when(
      Parameter_id %in% "SCB7" ~ "NS-EN ISO 6468",
      TRUE ~ "UKJENT"),
    Operator = case_when(
      is.na(FLAG1) ~ "=",
      TRUE ~ "<"),
    Provenr = as.character(NA),
    Tid_provetak = floor_date(SAMPLE_DATE, unit = "day") %>% as.character(),
    Ovre_dyp = NA,              # different from water
    Nedre_dyp = NA,             # different from water
    DybdeEnhet = "NULL",        # different from water
    Ant_verdier = NA,
    Kommentar = "",
    Deteksjonsgrense = NA,
    Kvantifiseringsgrense = NA
  )

check <- df_biota_vm_1 %>% 
  filter(is.na(Parameter_id)) %>% 
  distinct(NAME) %>% 
  select(NAME)

```

### Data for Vannmiljo 2

```{r}

# See above
# fn <- "Data_to_Vannmiljo/urban_vm/REG_water_2019-2021_kovirbv.xlsx"
# readxl::excel_sheets(fn)
# df_example <- readxl::read_excel(fn, sheet = "REG_water_2019-2021_kovirbv")

df_biota_vm_2 <- df_biota_vm_1 %>% 
  select(-Kommentar) %>% 
  left_join(lookup_kommentar, by = "Parameter_id")

```

### Add ID codes 

```{r}

df_biota_vm_2 <- df_biota_vm_2 %>%
  mutate(
    Vannlok_kode = paste0("NIVA@", STATION_ID),  # note: called 'Kilde_id' in station file (LOK)  
    ID_lokal = paste0("NIVA@BIO@", VALUE_ID)     # note: BIOTA version
  )

```


### Data for Vannmiljo 3
```{r}

# names(df_example) %>% dput()

vars <- c("ID_lokal", "Vannlok_kode", "Aktivitet_id", "Oppdragsgiver", 
          "Oppdragstaker", "Parameter_id", "Medium_id", "LatinskNavn_id", 
          "Provetakmetode_id", "Filtrert_prove", "Analysemetode_id", "Operator", 
          "Verdi", "Enhet_id", "Provenr", "Tid_provetak", "Ovre_dyp", "Nedre_dyp", 
          "DybdeEnhet", "Ant_verdier", "Kommentar", "Deteksjonsgrense", 
          "Kvantifiseringsgrense")

setdiff(vars, names(df_biota_vm_2))

df_vm_biota <- df_biota_vm_2[vars]


```

## 6. Combine water and biota data    

### Combine data  

```{r}

df_vm_combined <- bind_rows(df_vm_water, df_vm_biota)

```

### Units: All metals to mg/kg v.v. 

* All metals including mercury and rare earth metals should be in unit 42 (mg/kg v.v.) 

```{r}

if (FALSE){
  # check with Dag Rosland mail 
  # These have wrong units according to Dag R.:
  metals_to_change <- c("DY","ER","EU","GD","HG", "HO","LA","LU","ND","PR","SB","SC","SM","TB","TM","YB")  
  # Check that all are given as Group = metals:
  lookup_substances %>% 
    filter(ParameterID %in% metals_to_change) %>% 
    xtabs(~addNA(Group), .)
}

sel_rows <- df_biota_3

```

### Check parameters    

```{r}

check <- df_vm_combined %>% 
  filter(is.na(Parameter_id))
if (nrow(check) > 0){
  stop("some lacking!")
}

```

### Check stations
```{r}

check <- df_vm_combined %>% 
  filter(!Vannlok_kode %in% data_list$LOK$Kilde_id) %>% 
  distinct(Vannlok_kode)

if (nrow(check) > 0){
  stop("some lacking!")
}

```

### Check other columns    

```{r}

apply(is.na(df_vm_combined), 2, sum)

```

## 7. Final delivery  

### Make data 
```{r}

data_list <- list(
  LOK = df_stations_vm,
  REG = df_vm_combined
)


```




### Save
```{r}


t <- now(tzone = 'UTC') 
timestring <- paste0(substr(t, 1, 10), "_t", substr(t, 12, 13), substr(t, 15, 16))

fn <- glue::glue("Data_to_Vannmiljo/Data_{program}/DataVM_{program}_{sampling_year}_{timestring}.xlsx")
fn
writexl::write_xlsx(data_list, path = fn)

```

## APPENDIX: finding VANNMILJO tables in AqM

```{r}

get_nivabase_data(
"select * from ALL_TAB_COLUMNS 
where OWNER = 'VANNMILJO' and COLUMN_NAME = 'VANNKATEGORI';"
)

get_nivabase_data(
"select * from VANNMILJO.LOGG_LOKALITETER 
where rownum < 11"
)

get_nivabase_data(
"select * from VANNMILJO.LOGG_REGISTRERINGER 
where rownum < 11"
)


```
