---
title: "120_Create_2020data_for_Nivabase"
author: "DHJ"
date: "21 4 2022"
output: html_document
---

## 1. Setup  

### Packages     
```{r, message=FALSE, warning=FALSE, error=FALSE, results='hide'}

# detach("package:niRvana", unload=TRUE)

library(niRvana)
library(dplyr)
library(purrr)
library(lubridate)
library(readxl)
library(ggplot2)
library(tidyr)
library(fuzzyjoin)   # regex_full_join

# source("991_Vannmiljo_snippets_functions.R")
# source("992_Vannmiljo_Urban_fjord_functions.R")
# source("994_Industry_data_functions.R")         # for add_coordinates()

knitr::opts_chunk$set(results = 'hold')

```

## Stations from 2019 data

```{r}

df_stations <- read_excel("Input_data/Lookup_files/lookup_stations_urbanfjord.xlsx")

```


## Read data  

### NILU and IFE data   
* Exported from JMP as (note: decimal comma and utf-8 encoding)
```{r}

dat_nilu <- read.delim("Input_data/2020/NILU/004C NILU data stacked.txt", 
                       dec = ",", encoding = "UTF-8")

dat_ife <- read.delim("Input_data/2020/IFE/010 LIMS numbers and data IFE.txt", 
                      dec = ",", encoding = "UTF-8") %>%
  rename(LIMS.NR = LIMS_NR)


```


## Make lists of substancs  

```{r}

dat_nilu %>%
  count(Label) %>%
  writexl::write_xlsx("Input_data/2020/Substances_NILU_urbanfjord.xlsx")

```


## Check Labware stations      

### Check NILU data  
```{r}

# table(dat_nilu[,1]) # NILU number
# table(dat_nilu[,2]) # Matrix/species loose description 
# table(dat_nilu[,3]) # Localtion
table(dat_nilu[,4])  %>% head(25)

```

### Get samples data from LIMS      
* Based on NILU samples only   
* Uses LABWARE_IMPORT and LABWARE_CHECK_SAMPLE to get sample information   
```{r}

ids_all <- unique(dat_nilu$LIMS.NR)
sel <- grepl("NR", ids_all)
ids <- ids_all[sel]

cat("No LIMS number: \n")
ids_all[!sel]

# detach("package:niRvana", unload=TRUE)
# library(niRvana)

# debugonce(get_labware_samples)
df_samples_labware_raw <- get_labware_samples(ids)  

```

```{r}

dat_labware %>%
  distinct(ACCOUNT_NUMBER, ANALYSEOPPDRAG, SAMPLE_TYPE, DESCRIPTION, TEXT_ID, AQUAMONITOR_CODE)

```

## Overview tables

### Overview of samples that are in LIMS  
* Includes number of samples with NILU and IFE data    
* Also includes all 'AQUAMONITOR_CODE' (usually just one) and example of DESCRIPTION  
```{r}

df_samples_labware_raw %>%
  distinct(ANALYSEOPPDRAG, SAMPLE_TYPE, DESCRIPTION, TEXT_ID, AQUAMONITOR_CODE) %>%
  arrange(ANALYSEOPPDRAG, DESCRIPTION) %>%
  left_join(select(df_samples_labware_raw, TEXT_ID, BIOTA_SAMPLENO)) %>%
  left_join(dat_nilu %>% count(LIMS.NR, name = "NILU_nn"), by = c(TEXT_ID="LIMS.NR")) %>%
  left_join(dat_ife %>% count(LIMS.NR, name = "IFE_nn"), by = c(TEXT_ID="LIMS.NR")) %>%
  group_by(ANALYSEOPPDRAG, SAMPLE_TYPE, BIOTA_SAMPLENO) %>%
  summarise(NIVA_n = n(),
            NILU_n = sum(NILU_nn > 0), 
            IFE_n = sum(IFE_nn > 0), 
            AQUAMONITOR_CODE_all = paste(unique(AQUAMONITOR_CODE), collapse = ";"),
            DESCRIPTION_example = first(DESCRIPTION)
            )
  
``` 


```{r}

df <-  get_nivabase_selection(
    "ACCOUNT_NUMBER, ANALYSEOPPDRAG, SAMPLE_TYPE, DESCRIPTION, TEXT_ID, AQUAMONITOR_CODE", 
    "LABWARE_IMPORT", 
    "TEXT_ID", unique(dat_labware$TEXT_ID), values_are_text = TRUE,
    extra_sql = "")
nrow(df)


df <-  get_nivabase_selection(
    "ACCOUNT_NUMBER, ANALYSEOPPDRAG, SAMPLE_TYPE, DESCRIPTION, TEXT_ID, AQUAMONITOR_CODE, count(*)", 
    "LABWARE_IMPORT", 
    "TEXT_ID", unique(dat_labware$TEXT_ID), values_are_text = TRUE,
    extra_sql = " and REPORTED_NAME != 'Uspesifikk organisk analyse' group by ACCOUNT_NUMBER, ANALYSEOPPDRAG, SAMPLE_TYPE, DESCRIPTION, TEXT_ID, AQUAMONITOR_CODE")
nrow(df)

```


```{r}



```


### Add missing station ID in 'df_samples_labware'     

* Note: these will probably lack in Nivabasen (BIOTA_SAMPLES)  
* See message when running 'get_labware_samples' above: 

```
Warning in get_labware_samples(ids) :
  1 samples had no AQUAMONITOR_ID.  
DESCRIPTION content(s):Sediment-Cm21
```

```{r}

df_samples_labware <- df_samples_labware_raw

sel1 <- df_samples_labware$DESCRIPTION %in% "Sediment-Cm21"    # this is the one lacking station ID
sel2 <- with(df_stations, STATION_NAME == "Cm21" & SAMPLE_TYPE == "SEDIMENT")

if (sum(sel2) > 1){
  stop("Must select only one row")
} else {
  cat("STATION_ID:", df_stations$STATION_ID[sel2], "\n")
  cat("AQUAMONITOR_CODE:", df_stations$AQUAMONITOR_CODE[sel2])
}

df_samples_labware$AQUAMONITOR_ID[sel1] <- df_stations$STATION_ID[sel2]
df_samples_labware$AQUAMONITOR_CODE[sel1] <- df_stations$AQUAMONITOR_CODE[sel2]

```

### For hendelse

```
Prosjekt: O 17146 Urbanfjord, kontonummer 629

1. Biota-prøvene trenger prøvenummer (BIOTA_SAMPLENO), hentes fra DESCRIPTION. Nå står det bare null.
NR-2020-08402 - NR-2020-08431
NR-2020-08471 - NR-2020-08479
NR-2020-08481 - NR-2020-08515


2. NR-2020-08522 (beskrivelse Sediment-Cm21) skal ha
STATION_ID: 67067 
AQUAMONITOR_CODE: Cm21

```

```{r}

x <- df_samples_labware_raw %>%
  filter(SAMPLE_TYPE == "BIOTA") %>%
  arrange(TEXT_ID) %>%
  pull(TEXT_ID)

x2 <- substr(x, 9, 13) %>% as.numeric()

diff(x2)
x[diff(x2) > 1]

x

```


### Check whether all NILU data are found in Labware  
```{r}

check <- dat_nilu %>% 
  filter(!grepl("galle", LIMS.NR)) %>%
  filter(!LIMS.NR %in% df_samples_labware$TEXT_ID)

if (nrow(check) > 0){
  stop("Some samples that are not 'galle' are not found in Labware ")
} else {
  message("All samples that are not 'galle' are found in Labware")
}

```
### Check whether all IFE data are found in Labware  
```{r}

check <- dat_ife %>%                                # change no. 1 from previous chunk
  filter(!grepl("galle", LIMS.NR)) %>%
  filter(!grepl("blandprøve", Matrix_orig)) %>%   # change no. 2 from previous chunk
  filter(!LIMS.NR %in% df_samples_labware$TEXT_ID)

if (nrow(check) > 0){
  stop("Some samples that are not 'galle' are not found in Labware ")
} else {
  message("All samples that are not 'galle' are found in Labware")
}

```

## Check existing data in the database     

### In Nivabasen   
* No data found for 2020  
```{r}

# Get all specimens collected at these stations 
df_specimens <- get_specimens_from_stationdata(df_stations, years = 2020)
# NO SPECIMENS

df_specimens_2019 <- get_specimens_from_stationdata(df_stations, years = 2019)
# 45 specimens

# Get a list of projects
df_projects <- get_projects()   # we call it 'df_projects' (the default name)

# Get a list of stations
df_stations_from_project <- get_stations_from_project("urban", ignore.case = TRUE)

# Get all specimens, 2nd try
df_specimens <- get_specimens_from_stationdata(df_stations_from_project, years = 2020)

```


### 2020 data in Labware  
```{r}

dat_labware <- get_labware_import(df_samples_labware)

xtabs(~is.na(PARAM_I_AQUAMONITOR) + is.na(PARAM) + SAMPLE_TYPE, dat_labware)
xtabs(~is.na(PARAM_I_AQUAMONITOR) + is.na(REPORTED_NAME), dat_labware)

dat_labware %>%
  count(REPORTED_NAME, SAMPLE_TYPE, PARAM, PARAM_I_AQUAMONITOR) %>% View("Labware 2020")


```

```{r}

dat_labware %>%
  count(ACCOUNT_NUMBER, ANALYSEOPPDRAG, SAMPLE_TYPE)



dat_labware %>%
  arrange(ANALYSEOPPDRAG, DESCRIPTION) %>%
  filter(ANALYSEOPPDRAG %in% c("629-9460", "629-9461")) %>%
  distinct(TEXT_ID, DESCRIPTION) %>%
  count(DESCRIPTION)

dat_labware %>%
  filter(ANALYSEOPPDRAG %in% c("629-9460", "629-9461")) %>%
  xtabs(~REPORTED_NAME + addNA(ENTRY_QUALIFIER), .)

dat_labware %>%
  arrange(ANALYSEOPPDRAG, DESCRIPTION) %>%
  filter(ANALYSEOPPDRAG %in% c("629-9460", "629-9461") & ) %>%
  distinct(TEXT_ID, DESCRIPTION) %>%
  count(DESCRIPTION)

df_samples_labware_raw %>%
  count(SAMPLE_TYPE, SPECIES, TISSUE, BIOTA_SAMPLENO)

```

### Compare with 2019 biota data in Nivabasen    
```{r}

# ?get_biota_chemistry

df_specimens_2019 <- get_specimens_from_stationdata(df_stations, years = 2019)

# Get sample IDs
sampleids <- get_sampleid_from_specimendata(df_specimens_2019)

# Get data frame of samples
df_samples_2019 <- get_samples_from_sampleid(sampleids)

# Get data frame of chemical results (30 seconds or so)
dat_2019 <- get_biota_chemistry(
  years = 2019,
  specimendata = df_specimens_2019,
  stationdata = df_stations,
  report_samples = TRUE)

dat_2019 %>%
  count(REPORTED_NAME, SAMPLE_TYPE, PARAM, PARAM_I_AQUAMONITOR) %>% View("Labware 2020")


```



